---
title: "Simulation Study: Recovery Performance with Multiple Clustering Structures"
author: "Your Name"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, 
                      fig.width = 12, fig.height = 8, cache = TRUE)
```

# Executive Summary

This simulation study extends our analysis to investigate how the number of independent clustering structures affects recovery performance. We test whether the approach scales from 2 structures (our baseline) to 3, 4, and 5 independent structures, all embedded in 100-dimensional data.

**Key Questions**:
1. How does recovery performance degrade as the number of structures increases?
2. Is there a maximum number of structures that can be reliably recovered?
3. How do sample size requirements change with more structures?
4. Does the critical separation threshold shift with the number of structures?

# 1. Methods

## 1.1 Experimental Design

We use high-dimensional data (p=100) divided equally among k structures:
- **k=2**: 50 dimensions per structure (baseline)
- **k=3**: 33-34 dimensions per structure
- **k=4**: 25 dimensions per structure  
- **k=5**: 20 dimensions per structure

Each structure has a different number of clusters (3, 4, 5, 6, or 7) to ensure distinguishability.

### Experimental Grid:
- **Number of structures**: 2, 3, 4, 5
- **Dimensions**: Fixed at p=100
- **Sample sizes**: 250, 500, 1000
- **Separation factors**: 6.0σ, 3.0σ, 1.5σ, 0.5σ
- **Replications**: 25
- **Parallel cores**: 8

Total: 48 conditions × 25 replications = 900 runs

# 2. Implementation

```{r libraries}
# Load required libraries
library(mvtnorm)      # For multivariate normal data generation
library(cluster)      # For PAM clustering
library(mclust)       # For model-based clustering
library(ggplot2)      # For visualisation
library(dplyr)        # For data manipulation
library(tidyr)        # For data reshaping
library(viridis)      # For colour scales
library(gridExtra)    # For arranging plots
library(foreach)      # For parallel loops
library(doParallel)   # For parallel backend
library(parallel)     # For makeCluster and stopCluster

# Set seed for reproducibility
set.seed(42)
```

```{r parallel_setup}
# Note: Parallel cluster will be created within the simulation chunk
# to ensure proper scoping in R Markdown environment

cat("Parallel processing configuration:\n")
cat("- Total cores available:", detectCores(), "\n")
cat("- Cores to be used:", detectCores() - 2, "\n")
cat("- Cores kept free for system:", 2, "\n\n")
```

```{r cache_setup}
# Set up caching directory
cache_dir <- "multiple_structures_cache"
if (!dir.exists(cache_dir)) {
  dir.create(cache_dir)
  cat("Created cache directory:", cache_dir, "\n")
} else {
  cat("Using existing cache directory:", cache_dir, "\n")
}

# Cache functions
get_cache_filename <- function(n_structures, n_samples, separation_factor) {
  filename <- sprintf("results_k%d_n%d_sep%.1f.rds", n_structures, n_samples, separation_factor)
  file.path(cache_dir, filename)
}

check_condition_cache <- function(n_structures, n_samples, separation_factor) {
  cache_file <- get_cache_filename(n_structures, n_samples, separation_factor)
  file.exists(cache_file)
}

save_condition_results <- function(results, n_structures, n_samples, separation_factor) {
  cache_file <- get_cache_filename(n_structures, n_samples, separation_factor)
  saveRDS(results, cache_file)
}

load_condition_results <- function(n_structures, n_samples, separation_factor) {
  cache_file <- get_cache_filename(n_structures, n_samples, separation_factor)
  if (file.exists(cache_file)) {
    return(readRDS(cache_file))
  } else {
    return(NULL)
  }
}
```

```{r parameters}
# Experimental parameters
n_structures_values <- c(2, 3, 4, 5)
p_dims <- 100  # Fixed
sample_sizes <- c(250, 500, 1000)
separation_factors <- c(6, 3, 1.5)
n_replications <- 25

# Fixed parameters
cluster_variance <- 0.3

# Clusters per structure (different for each to ensure distinguishability)
clusters_per_structure <- c(3, 4, 5, 6, 7)  # Will use first k of these

# Calculate dimensions per structure for each k
dims_per_structure <- list(
  "2" = c(50, 50),
  "3" = c(34, 33, 33),
  "4" = c(25, 25, 25, 25),
  "5" = c(20, 20, 20, 20, 20)
)

# Total conditions
n_conditions <- length(n_structures_values) * length(sample_sizes) * length(separation_factors)

cat("\nExperimental design:\n")
cat("- Number of structures (k):", paste(n_structures_values, collapse=", "), "\n")
cat("- Total dimensions:", p_dims, "\n")
cat("- Sample sizes:", paste(sample_sizes, collapse=", "), "\n")
cat("- Separation factors:", paste(separation_factors, collapse=", "), "σ\n")
cat("- Total conditions:", n_conditions, "\n")
cat("- Replications per condition:", n_replications, "\n")
cat("- Total simulation runs:", n_conditions * n_replications, "\n")
```

```{r data_generation_functions}
# Generate cluster means for k structures
generate_cluster_means_k <- function(separation_factor, cluster_variance, n_structures, dims_list) {
  base_separation <- separation_factor * sqrt(cluster_variance)
  
  all_means <- list()
  
  for (k in 1:n_structures) {
    n_clusters <- clusters_per_structure[k]
    n_dims <- dims_list[k]
    
    # Generate means for this structure
    structure_means <- list()
    for (c in 1:n_clusters) {
      # Create distinct patterns for each cluster
      if (c <= 2) {
        # First two clusters: opposite sides
        multiplier <- ifelse(c == 1, -1, 1)
        means <- runif(n_dims, 0.8, 1.5) * multiplier * base_separation
      } else {
        # Additional clusters: random positions
        means <- runif(n_dims, -0.7, 0.7) * base_separation
        # Add some structure to avoid complete overlap
        means <- means + (c - 2.5) * 0.3 * base_separation
      }
      structure_means[[paste0("cluster_", c)]] <- means
    }
    
    all_means[[paste0("structure_", k)]] <- structure_means
  }
  
  return(all_means)
}

# Generate dataset with k structures
generate_dataset_k <- function(separation_factor, cluster_variance, n_samples, n_structures) {
  dims_list <- dims_per_structure[[as.character(n_structures)]]
  
  # Generate cluster assignments for each structure
  cluster_assignments <- list()
  for (k in 1:n_structures) {
    n_clusters <- clusters_per_structure[k]
    # Use different proportions for each structure
    props <- diff(c(0, sort(runif(n_clusters-1)), 1))
    cluster_assignments[[k]] <- sample(1:n_clusters, n_samples, replace = TRUE, prob = props)
  }
  
  # Get cluster means
  all_means <- generate_cluster_means_k(separation_factor, cluster_variance, n_structures, dims_list)
  
  # Initialize data matrix
  data_matrix <- matrix(NA, nrow = n_samples, ncol = p_dims)
  
  # Generate data for each structure
  col_start <- 1
  for (k in 1:n_structures) {
    n_dims <- dims_list[k]
    col_end <- col_start + n_dims - 1
    
    structure_means <- all_means[[paste0("structure_", k)]]
    
    # Generate data for this structure
    for (i in 1:n_samples) {
      cluster_id <- cluster_assignments[[k]][i]
      cluster_mean <- structure_means[[paste0("cluster_", cluster_id)]]
      cluster_cov <- diag(rep(cluster_variance, n_dims))
      
      data_matrix[i, col_start:col_end] <- rmvnorm(1, mean = cluster_mean, sigma = cluster_cov)
    }
    
    col_start <- col_end + 1
  }
  
  return(list(
    data = data_matrix,
    true_assignments = cluster_assignments,
    dims_list = dims_list
  ))
}
```

```{r analysis_functions}
# Analyze dataset with k structures
analyze_dataset_k <- function(data_matrix, n_structures, dims_list) {
  n_dims <- ncol(data_matrix)
  
  # Step 1: MCLUST on each dimension
  cluster_assignments <- matrix(NA, nrow = nrow(data_matrix), ncol = n_dims)
  detected_clusters <- numeric(n_dims)
  
  for(i in 1:n_dims) {
    mclust_fit <- Mclust(data_matrix[, i], G = 1:9, verbose = FALSE)
    cluster_assignments[, i] <- mclust_fit$classification
    detected_clusters[i] <- mclust_fit$G
  }
  
  # Step 2: Compute ARI matrix
  ari_matrix <- matrix(1, nrow = n_dims, ncol = n_dims)
  for(i in 1:(n_dims-1)) {
    for(j in (i+1):n_dims) {
      ari_val <- adjustedRandIndex(cluster_assignments[, i], cluster_assignments[, j])
      ari_matrix[i, j] <- ari_val
      ari_matrix[j, i] <- ari_val
    }
  }
  
  # Step 3: HCLust clustering
  dissim_matrix <- 1 - ari_matrix
  diag(dissim_matrix) <- 0
  dissim_dist <- as.dist(dissim_matrix)
  
  # Try k values around true number of structures
  min_k <- max(2, n_structures-1)
  max_k <- min(n_structures+2, 8)
  k_values <- min_k:max_k
  
    # Try Dunn index first
  opt <- tryCatch(
    NbClust::NbClust(diss = dissim_dist, distance = NULL,
                     min.nc = min_k, max.nc = max_k, method = "ward.D2",
                     index = "dunn"),
    error = function(e) NULL
  )
  
  # Initialize return variables
  index <- NA_character_
  optimal_k <- NA_integer_
  optimal_hclust <- NULL
  index_values <- NULL
  optimal_idx <- NA_integer_
  
  # Check if NbClust with Dunn index worked
  if (!is.null(opt) && !is.null(opt$Best.nc)) {
    best_k <- suppressWarnings(as.numeric(opt$Best.nc[1]))
    dunn_val <- suppressWarnings(as.numeric(opt$Best.nc[2]))
    
    if (!is.na(best_k) && best_k > 1 && !is.nan(dunn_val) && !is.infinite(dunn_val) && dunn_val > 0) {
      # Dunn index valid
      index <- "dunn"
      optimal_k <- best_k
      
      # Handle Best.partition (cluster assignment vector or matrix)
      if (is.matrix(opt$Best.partition)) {
        # Assume second row is actual partition (based on your previous logic)
        optimal_hclust <- as.vector(opt$Best.partition[2, ])
      } else {
        optimal_hclust <- as.vector(opt$Best.partition)
      }
      
      # Get Dunn index values for all tested cluster numbers
      index_values <- as.numeric(opt$All.index)
      all_index_names <- names(opt$All.index)
      optimal_idx <- which(all_index_names == as.character(best_k))
      if (length(optimal_idx) == 0) {
        optimal_idx <- 1
      }
      
    } else {
      # Dunn index invalid → fallback to silhouette
      index <- "silhouette"
      
      k_values <- min_k:max_k
      index_values <- numeric(length(k_values))
      hclust_results <- vector("list", length(k_values))
      
      for (i in seq_along(k_values)) {
        k <- k_values[i]
        hclust_fit <- hclust(dissim_dist, method = "ward.D2")
        clusters <- cutree(hclust_fit, k = k)
        hclust_results[[i]] <- clusters
        sil <- silhouette(clusters, dist = dissim_dist)
        index_values[i] <- mean(sil[, "sil_width"])
        cat(sprintf("\nk = %d: Average silhouette width = %.3f", k, index_values[i]))
      }
      
      optimal_idx <- which.max(index_values)
      optimal_k <- k_values[optimal_idx]
      optimal_hclust <- hclust_results[[optimal_idx]]
      
      cat("\nOptimal number of dimension groups (silhouette):", optimal_k, "\n")
    }
    
  } else {
    # NbClust failed (e.g., error or null), fallback to silhouette
    index <- "silhouette"
    
    k_values <- min_k:max_k
    index_values <- numeric(length(k_values))
    hclust_results <- vector("list", length(k_values))
    
    for (i in seq_along(k_values)) {
      k <- k_values[i]
      hclust_fit <- hclust(dissim_dist, method = "ward.D2")
      clusters <- cutree(hclust_fit, k = k)
      hclust_results[[i]] <- clusters
      sil <- silhouette(clusters, dist = dissim_dist)
      index_values[i] <- mean(sil[, "sil_width"])
      cat(sprintf("\nk = %d: Average silhouette width = %.3f", k, index_values[i]))
    }
    
    optimal_idx <- which.max(index_values)
    optimal_k <- k_values[optimal_idx]
    optimal_hclust <- hclust_results[[optimal_idx]]
    
    cat("\nOptimal number of dimension groups (silhouette):", optimal_k, "\n")
  }
  
  # Step 4: Calculate recovery metrics
  # Create true structure labels
  true_labels <- rep(NA, n_dims)
  col_start <- 1
  for (k in 1:n_structures) {
    col_end <- col_start + dims_list[k] - 1
    true_labels[col_start:col_end] <- k
    col_start <- col_end + 1
  }
  
  recovery_ari <- adjustedRandIndex(optimal_hclust, true_labels)
  
  # Check detection accuracy for each structure
  structure_detection <- numeric(n_structures)
  col_start <- 1
  for (k in 1:n_structures) {
    col_end <- col_start + dims_list[k] - 1
    expected_clusters <- clusters_per_structure[k]
    structure_detection[k] <- sum(detected_clusters[col_start:col_end] == expected_clusters) / dims_list[k]
    col_start <- col_end + 1
  }
  
  return(list(
    recovery_ari = recovery_ari,
    optimal_k = optimal_k,
    found_correct_k = (optimal_k == n_structures),
    index = index,
    index_value = as.numeric(index_values[optimal_idx]),
    structure_detection = structure_detection,
    mean_detection = mean(structure_detection)
  ))
}

# Wrapper function for parallel execution
run_single_replication <- function(rep, n_structures, separation_factor, n_samples, dims_list) {
  # Generate dataset
  dataset <- generate_dataset_k(separation_factor, cluster_variance, n_samples, n_structures)
  
  # Analyze dataset
  analysis_result <- analyze_dataset_k(dataset$data, n_structures, dims_list)
  
  # Return row of results
  data.frame(
    n_structures = n_structures,
    n_samples = n_samples,
    separation_factor = separation_factor,
    replication = rep,
    recovery_ari = analysis_result$recovery_ari,
    optimal_k = analysis_result$optimal_k,
    found_correct_k = analysis_result$found_correct_k,
    index = analysis_result$index,
    index_value = analysis_result$index_value,
    mean_detection = analysis_result$mean_detection,
    stringsAsFactors = FALSE
  )
}
```

# 3. Run Simulation

```{r run_simulation, cache=FALSE}
# Set up parallel processing within this chunk
n_cores <- parallel::detectCores() - 2  # Leave 2 cores free for system
cl <- parallel::makeCluster(n_cores)
doParallel::registerDoParallel(cl)

# Initialize results storage
all_results <- data.frame()
condition_counter <- 0
start_time <- Sys.time()

cat("\nStarting parallel simulation with multiple structures...\n")
cat("Using", n_cores, "parallel cores (out of", parallel::detectCores(), "available)\n")
cat("Time started:", format(start_time, "%Y-%m-%d %H:%M:%S"), "\n\n")

# Loop through experimental conditions
for (n_structures in n_structures_values) {
  cat(sprintf("\n=== Processing k = %d structures ===\n", n_structures))
  structure_start_time <- Sys.time()
  
  dims_list <- dims_per_structure[[as.character(n_structures)]]
  
  for (n_samples in sample_sizes) {
    for (sep_factor in separation_factors) {
      condition_counter <- condition_counter + 1
      
      cat(sprintf("Condition %d/%d: k=%d, n=%d, sep=%.1fσ", 
                  condition_counter, n_conditions, n_structures, n_samples, sep_factor))
      
      # Check cache
      if (check_condition_cache(n_structures, n_samples, sep_factor)) {
        cat(" [Cached]\n")
        condition_results <- load_condition_results(n_structures, n_samples, sep_factor)
        
      } else {
        cat(" [Computing in parallel...]\n")
        
        # Run replications in parallel
        condition_results <- foreach(
          rep = 1:n_replications,
          .combine = rbind,
          .packages = c("mvtnorm", "mclust", "cluster", "otrimle"),
          .export = c("run_single_replication", "generate_dataset_k", "analyze_dataset_k",  
                      "generate_cluster_means_k", "dims_per_structure", "clusters_per_structure",  
                      "cluster_variance", "p_dims")
        ) %dopar% {
          run_single_replication(rep, n_structures, separation_factor = sep_factor, n_samples, dims_list)
        }

        
        # Cache results
        save_condition_results(condition_results, n_structures, n_samples, sep_factor)
      }
      
      # Add to overall results
      all_results <- rbind(all_results, condition_results)
    }
  }
  
  structure_end_time <- Sys.time()
  structure_duration <- difftime(structure_end_time, structure_start_time, units = "mins")
  cat(sprintf("Completed k=%d in %.1f minutes\n", n_structures, structure_duration))
}

end_time <- Sys.time()
total_duration <- difftime(end_time, start_time, units = "mins")

cat("\n=== SIMULATION COMPLETE ===\n")
cat("Total time:", sprintf("%.1f minutes (%.2f hours)\n", total_duration, total_duration/60))
cat("Total results collected:", nrow(all_results), "\n")

# Clean up parallel cluster with error handling
tryCatch({
  parallel::stopCluster(cl)
  cat("Parallel cluster stopped\n")
}, error = function(e) {
  cat("Note: Parallel cluster already stopped or disconnected\n")
})
```

# 4. Results

```{r join-results, echo=FALSE, message=FALSE, warning=FALSE}
# Set the path to your folder
folder_path <- "multiple_structures_cache"

# List all RDS files in the folder
file_list <- list.files(path = folder_path, pattern = "^results.*\\.rds$", full.names = TRUE)

# Read all RDS files and store them in a list
data_list <- lapply(file_list, readRDS)

# Assuming each RDS contains a data frame, combine them
all_results_ms <- do.call(rbind, data_list)

# Save the combined results into the same folder
file_path <- file.path(folder_path, "all_results_ms.rds")
saveRDS(all_results_ms, file_path)
```

## 4.1 Summary Statistics

```{r rename, echo=FALSE, message=FALSE, warning=FALSE}
# Set the path to your folder
all_results <- all_results_ms
```
```{r summary_stats}
# Calculate summary statistics
summary_stats <- all_results %>%
  group_by(n_structures, n_samples, separation_factor) %>%
  summarise(
    mean_ari = mean(recovery_ari),
    sd_ari = sd(recovery_ari),
    prop_perfect = mean(recovery_ari > 0.95),
    prop_failed = mean(recovery_ari < 0.5),
    prop_found_k = mean(found_correct_k),
    mean_k = mean(optimal_k),
    mean_index = mean(index_value),
    mean_detection = mean(mean_detection),
    .groups = 'drop'
  )

# Create summary table by number of structures
for (k in n_structures_values) {
  cat(sprintf("\n### Recovery Performance for k = %d structures\n", k))
  
  k_summary <- summary_stats %>%
    filter(n_structures == k) %>%
    select(n_samples, separation_factor, mean_ari, prop_found_k) %>%
    pivot_wider(names_from = separation_factor, 
                values_from = c(mean_ari, prop_found_k),
                names_sep = "_")
  
  # Format for display
display_table <- k_summary %>%
  mutate(
    `6.0σ`  = sprintf("%.3f (%.0f%%)", unlist(mean_ari_6),  unlist(prop_found_k_6) * 100),
    `3.0σ`  = sprintf("%.3f (%.0f%%)", unlist(mean_ari_3),  unlist(prop_found_k_3) * 100),
    `1.5σ`  = sprintf("%.3f (%.0f%%)", unlist(mean_ari_1.5), unlist(prop_found_k_1.5) * 100)
  ) %>%
  select(n_samples, `6.0σ`, `3.0σ`, `1.5σ`)

  
  print(knitr::kable(display_table,
                     caption = sprintf("Mean ARI (Proportion finding correct k) for k=%d", k),
                     col.names = c("Sample Size", "6.0σ", "3.0σ", "1.5σ")))
}
```

## 4.2 Main Visualisation: Effect of Number of Structures

```{r main_plot, fig.height=10}
summary_stats2 <- summary_stats %>% 
  mutate(n_samples_f = factor(n_samples))

# Create faceted plot by separation
p_main <- ggplot(summary_stats2, aes(x = n_structures, y = mean_ari, 
                                   colour = n_samples_f, 
                                   group = n_samples)) +
  geom_line(size = 1.2) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = mean_ari - sd_ari, ymax = mean_ari + sd_ari), 
                width = 0.1, alpha = 0.5) +
  facet_wrap(~ separation_factor, ncol = 2,
             labeller = labeller(separation_factor = function(x) paste(x, "σ separation"))) +
  scale_colour_viridis_d(name = "Sample\nSize", option = "viridis") +
  scale_x_continuous(breaks = n_structures_values) +
  scale_y_continuous(limits = c(0, 1.1), breaks = seq(0, 1, 0.2)) +
  geom_hline(yintercept = 0.8, linetype = "dashed", alpha = 0.5) +
  labs(title = "Recovery Performance vs Number of Structures",
       subtitle = "How does performance degrade as we add more independent clustering structures?",
       x = "Number of Structures (k)",
       y = "Mean Recovery ARI (± SD)") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, size = 12, face = "bold", family = "Times"),
        plot.subtitle = element_text(hjust = 0.5, size = 10, family = "Times"),
        axis.text = element_text(size = 10, family = "Times"),
        axis.title = element_text(size = 12, face = "bold", family = "Times"),
        legend.title = element_text(size = 12, face = "bold", family = "Times"),
        legend.text = element_text(size = 10, family = "Times"),
       strip.text = element_text(size = 12, family = "Times", face = "bold"),
        panel.spacing = unit(1, "lines"))

print(p_main)

ggsave(paste0("1j_NSrecovery.pdf"), plot = p_main, device = cairo_pdf, width = 6, height = 5, units = "in")
```

## 4.3 Heatmap Visualisation

```{r heatmap}
# Create heatmap
p_heatmap <- ggplot(summary_stats, aes(x = factor(n_samples), 
                                       y = factor(n_structures), 
                                       fill = mean_ari)) +
  geom_tile(colour = "white", size = 1) +
  geom_text(aes(label = sprintf("%.2f", mean_ari)), 
            colour = ifelse(summary_stats$mean_ari > 0.5, "white", "black"),
            size = 4) +
  facet_wrap(~ separation_factor, ncol = 3,
             labeller = labeller(separation_factor = function(x) paste(x, "σ separation"))) +
  scale_fill_viridis(name = "Mean\nRecovery\nARI", 
                     limits = c(0, 1),
                     option = "plasma") +
  scale_x_discrete(labels = format) +
  scale_y_discrete(labels = function(x) paste(x, "structures")) +
  labs(title = "Recovery Performance Heatmap",
       subtitle = "Mean ARI across different conditions",
       x = "Sample Size",
       y = "Number of Structures") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
        plot.subtitle = element_text(hjust = 0.5, size = 12),
        strip.text = element_text(size = 12, face = "bold"),
        panel.spacing = unit(1, "lines"))

print(p_heatmap)
```

## 4.4 Finding the Correct Number of Structures

```{r k_detection}
# Plot proportion finding correct k
p_k_detection <- ggplot(summary_stats, aes(x = n_structures, y = prop_found_k, 
                                          colour = factor(n_samples),
                                          group = n_samples)) +
  geom_line(size = 1.2) +
  geom_point(size = 3) +
  facet_wrap(~ separation_factor, ncol = 3,
             labeller = labeller(separation_factor = function(x) paste(x, "σ separation"))) +
  scale_colour_viridis_d(name = "Sample\nSize", option = "viridis") +
  scale_x_continuous(breaks = n_structures_values) +
  scale_y_continuous(limits = c(0, 1.05), breaks = seq(0, 1, 0.2)) +
  labs(title = "Accuracy in Detecting the Correct Number of Structures",
       subtitle = "Proportion of runs where HClust found k structures",
       x = "True Number of Structures (k)",
       y = "Proportion Finding Correct k") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
        plot.subtitle = element_text(hjust = 0.5, size = 12),
        strip.text = element_text(size = 12, face = "bold"),
        legend.position = "right")

print(p_k_detection)
```

## 4.5 Dimensions per Structure Analysis

```{r dims_analysis}
# Add dimensions per structure info
summary_with_dims <- summary_stats %>%
  mutate(dims_per_structure = p_dims / n_structures,
         samples_per_dim_per_struct = n_samples / dims_per_structure)

# Plot by dimensions per structure
p_dims <- ggplot(summary_with_dims, aes(x = dims_per_structure, y = mean_ari,
                                        colour = factor(separation_factor))) +
  geom_line(aes(group = interaction(n_samples, separation_factor)),
            alpha = 0.3) +
  geom_point(aes(shape = factor(n_samples)), size = 3) +
  scale_x_continuous(trans = "log10", 
                     breaks = c(20, 25, 33, 50),
                     labels = c(20, 25, 33, 50)) +
  scale_colour_viridis_d(name = "Separation", 
                        labels = function(x) paste0(x, "σ"),
                        option = "plasma") +
  scale_shape_manual(name = "Sample\nSize",
                     values = c(16, 17, 18)) +
  geom_hline(yintercept = 0.8, linetype = "dashed", alpha = 0.5) +
  labs(title = "Effect of Dimensions per Structure",
       subtitle = "Performance vs dimensions allocated to each structure (p=100 total)",
       x = "Dimensions per Structure (log scale)",
       y = "Mean Recovery ARI") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
        plot.subtitle = element_text(hjust = 0.5, size = 12))

print(p_dims)
```

