---
title: "Simulation Study: Recovery Performance with Unequal Structure Sizes"
author: "Your Name"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, 
                      fig.width = 12, fig.height = 8, cache = TRUE)
```

# Executive Summary

This simulation study extends our multiple structures analysis to investigate how **unequal allocation of dimensions** affects recovery performance. Instead of dividing dimensions equally among structures, we test scenarios where some structures dominate the dimensional space while others have minimal representation.

**Key Questions**:
1. Does having one dominant structure mask the detection of smaller structures?
2. Is there a minimum dimension threshold below which structures become undetectable?
3. How do unequal allocations compare to equal allocations in terms of recovery performance?
4. What allocation patterns are most challenging for the method?

# 1. Methods

## 1.1 Experimental Design

We test various allocation patterns for p=100 dimensions across k structures:

### Allocation Patterns:
- **Equal**: Baseline equal division (as before)
- **Dominant**: One structure gets 50-70% of dimensions
- **Gradual**: Decreasing allocation (e.g., 40%, 30%, 20%, 10%)
- **Extreme**: Very unbalanced (e.g., 80%, 10%, 10%)

### Fixed Parameters:
- Total dimensions: p = 100
- Number of structures: k = 2, 3, 4
- Sample sizes: 250, 500, 1000
- Separation: 2.0σ (where interesting transitions occur)
- Replications: 25

# 2. Implementation

```{r libraries}
# Load required libraries
library(mvtnorm)      # For multivariate normal data generation
library(cluster)      # For PAM clustering
library(mclust)       # For model-based clustering
library(ggplot2)      # For visualisation
library(dplyr)        # For data manipulation
library(tidyr)        # For data reshaping
library(viridis)      # For colour scales
library(gridExtra)    # For arranging plots
library(foreach)      # For parallel loops
library(doParallel)   # For parallel backend
library(parallel)     # For cluster management

# Set seed for reproducibility
set.seed(42)
```

```{r cache_setup}
# Set up caching directory
cache_dir <- "unequal_structures_cache"
if (!dir.exists(cache_dir)) {
  dir.create(cache_dir)
  cat("Created cache directory:", cache_dir, "\n")
} else {
  cat("Using existing cache directory:", cache_dir, "\n")
}

# Cache functions
get_cache_filename <- function(n_structures, allocation_type, n_samples) {
  filename <- sprintf("results_k%d_%s_n%d.rds", n_structures, allocation_type, n_samples)
  file.path(cache_dir, filename)
}

check_condition_cache <- function(n_structures, allocation_type, n_samples) {
  cache_file <- get_cache_filename(n_structures, allocation_type, n_samples)
  file.exists(cache_file)
}

save_condition_results <- function(results, n_structures, allocation_type, n_samples) {
  cache_file <- get_cache_filename(n_structures, allocation_type, n_samples)
  saveRDS(results, cache_file)
}

load_condition_results <- function(n_structures, allocation_type, n_samples) {
  cache_file <- get_cache_filename(n_structures, allocation_type, n_samples)
  if (file.exists(cache_file)) {
    return(readRDS(cache_file))
  } else {
    return(NULL)
  }
}
```

```{r parameters}
# Experimental parameters
n_structures_values <- c(2, 3, 4)
p_dims <- 100
sample_sizes <- c(250, 500, 1000)
separation_factor <- 2.0  # Fixed at interesting transition point
n_replications <- 25

# Fixed parameters
cluster_variance <- 0.3
clusters_per_structure <- c(3, 4, 5, 6)  # Different number for each structure

# Define allocation patterns
allocation_patterns <- list(
  "k2" = list(
    equal = c(50, 50),
    dominant = c(70, 30),
    extreme = c(85, 15)
  ),
  "k3" = list(
    equal = c(34, 33, 33),
    dominant = c(50, 30, 20),
    gradual = c(45, 35, 20),
    extreme = c(70, 20, 10)
  ),
  "k4" = list(
    equal = c(25, 25, 25, 25),
    dominant = c(40, 25, 20, 15),
    gradual = c(40, 30, 20, 10),
    extreme = c(55, 25, 15, 5)
  )
)

# Count total conditions
total_conditions <- 0
for (k in n_structures_values) {
  total_conditions <- total_conditions + 
    length(allocation_patterns[[paste0("k", k)]]) * length(sample_sizes)
}

cat("\nExperimental design:\n")
cat("- Number of structures (k):", paste(n_structures_values, collapse=", "), "\n")
cat("- Total dimensions:", p_dims, "\n")
cat("- Sample sizes:", paste(sample_sizes, collapse=", "), "\n")
cat("- Separation factor:", separation_factor, "σ\n")
cat("- Total conditions:", total_conditions, "\n")
cat("- Replications per condition:", n_replications, "\n")
cat("- Total simulation runs:", total_conditions * n_replications, "\n\n")

# Display allocation patterns
cat("Allocation patterns:\n")
for (k in n_structures_values) {
  cat(sprintf("\nk=%d structures:\n", k))
  patterns <- allocation_patterns[[paste0("k", k)]]
  for (pattern_name in names(patterns)) {
    cat(sprintf("  %s: %s\n", pattern_name, 
                paste(patterns[[pattern_name]], collapse=", ")))
  }
}
```

```{r data_generation_functions}
# Generate cluster means for k structures with specified dimension allocation
generate_cluster_means_unequal <- function(separation_factor, cluster_variance, 
                                          n_structures, dims_list) {
  base_separation <- separation_factor * sqrt(cluster_variance)
  
  all_means <- list()
  
  for (k in 1:n_structures) {
    n_clusters <- clusters_per_structure[k]
    n_dims <- dims_list[k]
    
    # Skip if structure has too few dimensions
    if (n_dims < 2) {
      warning(sprintf("Structure %d has only %d dimensions, skipping", k, n_dims))
      next
    }
    
    # Generate means for this structure
    structure_means <- list()
    for (c in 1:n_clusters) {
      if (c <= 2) {
        # First two clusters: opposite sides
        multiplier <- ifelse(c == 1, -1, 1)
        means <- runif(n_dims, 0.8, 1.5) * multiplier * base_separation
      } else {
        # Additional clusters: random positions with structure
        means <- runif(n_dims, -0.7, 0.7) * base_separation
        means <- means + (c - 2.5) * 0.3 * base_separation
      }
      structure_means[[paste0("cluster_", c)]] <- means
    }
    
    all_means[[paste0("structure_", k)]] <- structure_means
  }
  
  return(all_means)
}

# Generate dataset with unequal structure sizes
generate_dataset_unequal <- function(separation_factor, cluster_variance, n_samples, 
                                    n_structures, dims_list) {
  
  # Validate dimensions sum to p_dims
  if (sum(dims_list) != p_dims) {
    stop("Dimension allocations must sum to p_dims")
  }
  
  # Generate cluster assignments for each structure
  cluster_assignments <- list()
  for (k in 1:n_structures) {
    n_clusters <- clusters_per_structure[k]
    # Random proportions for each structure
    props <- diff(c(0, sort(runif(n_clusters-1)), 1))
    cluster_assignments[[k]] <- sample(1:n_clusters, n_samples, replace = TRUE, prob = props)
  }
  
  # Get cluster means
  all_means <- generate_cluster_means_unequal(separation_factor, cluster_variance, 
                                              n_structures, dims_list)
  
  # Initialize data matrix
  data_matrix <- matrix(NA, nrow = n_samples, ncol = p_dims)
  
  # Generate data for each structure
  col_start <- 1
  for (k in 1:n_structures) {
    n_dims <- dims_list[k]
    
    # Skip if structure has too few dimensions
    if (n_dims < 2) {
      col_start <- col_start + n_dims
      next
    }
    
    col_end <- col_start + n_dims - 1
    structure_means <- all_means[[paste0("structure_", k)]]
    
    # Generate data for this structure
    for (i in 1:n_samples) {
      cluster_id <- cluster_assignments[[k]][i]
      cluster_mean <- structure_means[[paste0("cluster_", cluster_id)]]
      cluster_cov <- diag(rep(cluster_variance, n_dims))
      
      data_matrix[i, col_start:col_end] <- rmvnorm(1, mean = cluster_mean, sigma = cluster_cov)
    }
    
    col_start <- col_end + 1
  }
  
  return(list(
    data = data_matrix,
    true_assignments = cluster_assignments,
    dims_list = dims_list
  ))
}
```

```{r analysis_functions}
# Analyze dataset with unequal structures
analyze_dataset_unequal <- function(data_matrix, n_structures, dims_list) {
  n_dims <- ncol(data_matrix)
  
  # Step 1: MCLUST on each dimension
  cluster_assignments <- matrix(NA, nrow = nrow(data_matrix), ncol = n_dims)
  detected_clusters <- numeric(n_dims)
  
  for(i in 1:n_dims) {
    mclust_fit <- Mclust(data_matrix[, i], G = 1:9, verbose = FALSE)
    cluster_assignments[, i] <- mclust_fit$classification
    detected_clusters[i] <- mclust_fit$G
  }
  
  # Step 2: Compute ARI matrix
  ari_matrix <- matrix(1, nrow = n_dims, ncol = n_dims)
  for(i in 1:(n_dims-1)) {
    for(j in (i+1):n_dims) {
      ari_val <- adjustedRandIndex(cluster_assignments[, i], cluster_assignments[, j])
      ari_matrix[i, j] <- ari_val
      ari_matrix[j, i] <- ari_val
    }
  }
  
  # Step 3: HClust clustering
  dissim_matrix <- 1 - ari_matrix
  diag(dissim_matrix) <- 0
  dissim_dist <- as.dist(dissim_matrix)
  
 # Try k values around true number of structures
  min_k <- max(2, n_structures-1)
  max_k <- min(n_structures+2, 8)
  k_values <- min_k:max_k
  
    # Try Dunn index first
  opt <- tryCatch(
    NbClust::NbClust(diss = dissim_dist, distance = NULL,
                     min.nc = min_k, max.nc = max_k, method = "ward.D2",
                     index = "dunn"),
    error = function(e) NULL
  )
  
  # Initialize return variables
  index <- NA_character_
  optimal_k <- NA_integer_
  optimal_hclust <- NULL
  index_values <- NULL
  optimal_idx <- NA_integer_
  
  # Check if NbClust with Dunn index worked
  if (!is.null(opt) && !is.null(opt$Best.nc)) {
    best_k <- suppressWarnings(as.numeric(opt$Best.nc[1]))
    dunn_val <- suppressWarnings(as.numeric(opt$Best.nc[2]))
    
    if (!is.na(best_k) && best_k > 1 && !is.nan(dunn_val) && !is.infinite(dunn_val) && dunn_val > 0) {
      # Dunn index valid
      index <- "dunn"
      optimal_k <- best_k
      
      # Handle Best.partition (cluster assignment vector or matrix)
      if (is.matrix(opt$Best.partition)) {
        # Assume second row is actual partition (based on your previous logic)
        optimal_hclust <- as.vector(opt$Best.partition[2, ])
      } else {
        optimal_hclust <- as.vector(opt$Best.partition)
      }
      
      # Get Dunn index values for all tested cluster numbers
      index_values <- as.numeric(opt$All.index)
      all_index_names <- names(opt$All.index)
      optimal_idx <- which(all_index_names == as.character(best_k))
      if (length(optimal_idx) == 0) {
        optimal_idx <- 1
      }
      
    } else {
      # Dunn index invalid → fallback to silhouette
      index <- "silhouette"
      
      k_values <- min_k:max_k
      index_values <- numeric(length(k_values))
      hclust_results <- vector("list", length(k_values))
      
      for (i in seq_along(k_values)) {
        k <- k_values[i]
        hclust_fit <- hclust(dissim_dist, method = "ward.D2")
        clusters <- cutree(hclust_fit, k = k)
        hclust_results[[i]] <- clusters
        sil <- silhouette(clusters, dist = dissim_dist)
        index_values[i] <- mean(sil[, "sil_width"])
        cat(sprintf("\nk = %d: Average silhouette width = %.3f", k, index_values[i]))
      }
      
      optimal_idx <- which.max(index_values)
      optimal_k <- k_values[optimal_idx]
      optimal_hclust <- hclust_results[[optimal_idx]]
      
      cat("\nOptimal number of dimension groups (silhouette):", optimal_k, "\n")
    }
    
  } else {
    # NbClust failed (e.g., error or null), fallback to silhouette
    index <- "silhouette"
    
    k_values <- min_k:max_k
    index_values <- numeric(length(k_values))
    hclust_results <- vector("list", length(k_values))
    
    for (i in seq_along(k_values)) {
      k <- k_values[i]
      hclust_fit <- hclust(dissim_dist, method = "ward.D2")
      clusters <- cutree(hclust_fit, k = k)
      hclust_results[[i]] <- clusters
      sil <- silhouette(clusters, dist = dissim_dist)
      index_values[i] <- mean(sil[, "sil_width"])
      cat(sprintf("\nk = %d: Average silhouette width = %.3f", k, index_values[i]))
    }
    
    optimal_idx <- which.max(index_values)
    optimal_k <- k_values[optimal_idx]
    optimal_hclust <- hclust_results[[optimal_idx]]
    
    cat("\nOptimal number of dimension groups (silhouette):", optimal_k, "\n")
  }
  
  # Step 4: Calculate recovery metrics
  # Create true structure labels
  true_labels <- rep(NA, n_dims)
  col_start <- 1
  for (k in 1:n_structures) {
    col_end <- col_start + dims_list[k] - 1
    true_labels[col_start:col_end] <- k
    col_start <- col_end + 1
  }
  
  recovery_ari <- adjustedRandIndex(optimal_hclust, true_labels)
  
  # Check recovery for each structure separately
  structure_recovery <- numeric(n_structures)
  col_start <- 1
  for (k in 1:n_structures) {
    if (dims_list[k] < 2) {
      structure_recovery[k] <- NA
      col_start <- col_start + dims_list[k]
      next
    }
    
    col_end <- col_start + dims_list[k] - 1
    structure_dims <- col_start:col_end
    
    # Check if structure k dimensions are clustered together
    structure_clusters <- optimal_hclust[structure_dims]
    most_common_cluster <- as.numeric(names(sort(table(structure_clusters), 
                                                decreasing = TRUE)[1]))
    structure_recovery[k] <- sum(structure_clusters == most_common_cluster) / dims_list[k]
    
    col_start <- col_end + 1
  }
  
  return(list(
    recovery_ari = recovery_ari,
    optimal_k = optimal_k,
    found_correct_k = (optimal_k == n_structures),
    index = index,
    index_value = as.numeric(index_values[optimal_idx]),
    structure_recovery = structure_recovery,
    mean_structure_recovery = mean(structure_recovery, na.rm = TRUE)
  ))
}

# Wrapper function for parallel execution
run_single_replication_unequal <- function(rep, n_structures, n_samples, 
                                          dims_list, separation_factor, allocation_type) {
  # Generate dataset
  dataset <- generate_dataset_unequal(separation_factor, cluster_variance, 
                                     n_samples, n_structures, dims_list)
  
  # Analyze dataset
  analysis_result <- analyze_dataset_unequal(dataset$data, n_structures, dims_list)
  
  # Create base results
  result_df <- data.frame(
    n_structures = n_structures,
    allocation_type = allocation_type,
    n_samples = n_samples,
    replication = rep,
    recovery_ari = analysis_result$recovery_ari,
    optimal_k = analysis_result$optimal_k,
    found_correct_k = analysis_result$found_correct_k,
    index = analysis_result$index,
    index_value = analysis_result$index_value,
    mean_structure_recovery = analysis_result$mean_structure_recovery,
    stringsAsFactors = FALSE
  )
  
  # Add structure-specific results as separate columns
  # Use fixed column names that work for all k values
  for (s in 1:4) {  # Max 4 structures in our design
    if (s <= n_structures) {
      result_df[[paste0("struct", s, "_recovery")]] <- analysis_result$structure_recovery[s]
      result_df[[paste0("struct", s, "_dims")]] <- dims_list[s]
    } else {
      result_df[[paste0("struct", s, "_recovery")]] <- NA
      result_df[[paste0("struct", s, "_dims")]] <- NA
    }
  }
  
  return(result_df)
}
```

# 3. Run Simulation

```{r run_simulation, cache=FALSE}
# Set up parallel processing
n_cores <- parallel::detectCores() - 2
cl <- parallel::makeCluster(n_cores)
doParallel::registerDoParallel(cl)

# Initialize results storage
all_results <- data.frame()
condition_counter <- 0
start_time <- Sys.time()

cat("\nStarting parallel simulation with unequal structure sizes...\n")
cat("Using", n_cores, "parallel cores\n")
cat("Time started:", format(start_time, "%Y-%m-%d %H:%M:%S"), "\n\n")

# Loop through experimental conditions
for (n_structures in n_structures_values) {
  cat(sprintf("\n=== Processing k = %d structures ===\n", n_structures))
  structure_start_time <- Sys.time()
  
  # Get allocation patterns for this k
  patterns <- allocation_patterns[[paste0("k", n_structures)]]
  
  for (pattern_name in names(patterns)) {
    dims_list <- patterns[[pattern_name]]
    
    for (n_samples in sample_sizes) {
      condition_counter <- condition_counter + 1
      
      cat(sprintf("Condition %d/%d: k=%d, pattern=%s, n=%d", 
                  condition_counter, total_conditions, 
                  n_structures, pattern_name, n_samples))
      cat(sprintf(" [dims: %s]", paste(dims_list, collapse=",")))
      
      # Check cache
      if (check_condition_cache(n_structures, pattern_name, n_samples)) {
        cat(" [Cached]\n")
        condition_results <- load_condition_results(n_structures, pattern_name, n_samples)
        
      } else {
        cat(" [Computing...]\n")
        
        # Run replications in parallel
        condition_results <- foreach(
          rep = 1:n_replications,
          .combine = rbind,
          .packages = c("mvtnorm", "mclust", "cluster"),
          .export = c("run_single_replication_unequal", "generate_dataset_unequal",
                     "analyze_dataset_unequal", "generate_cluster_means_unequal",
                     "clusters_per_structure", "cluster_variance", "p_dims",
                     "separation_factor")) %dopar% {
                       
          run_single_replication_unequal(rep, n_structures, n_samples, 
                                        dims_list, separation_factor, pattern_name)
        }

        # Cache results
        save_condition_results(condition_results, n_structures, pattern_name, n_samples)
      
      }
      # Add to overall results
      all_results <- rbind(all_results, condition_results)
    }
  }
  
  structure_end_time <- Sys.time()
  structure_duration <- difftime(structure_end_time, structure_start_time, units = "mins")
  cat(sprintf("Completed k=%d in %.1f minutes\n", n_structures, structure_duration))
}

end_time <- Sys.time()
total_duration <- difftime(end_time, start_time, units = "mins")

cat("\n=== SIMULATION COMPLETE ===\n")
cat("Total time:", sprintf("%.1f minutes (%.2f hours)\n", total_duration, total_duration/60))
cat("Total results collected:", nrow(all_results), "\n")

# Clean up parallel cluster
tryCatch({
  parallel::stopCluster(cl)
  cat("Parallel cluster stopped\n")
}, error = function(e) {
  cat("Note: Parallel cluster already stopped or disconnected\n")
})
```

# 4. Results

## 4.1 Summary Statistics

```{r join-results, echo=FALSE, message=FALSE, warning=FALSE}
# Set the path to your folder
folder_path <- "unequal_structures_cache"

# List all RDS files in the folder
file_list <- list.files(path = folder_path, pattern = "^results.*\\.rds$", full.names = TRUE)

# Read all RDS files and store them in a list
data_list <- lapply(file_list, readRDS)

# Assuming each RDS contains a data frame, combine them
all_results_us <- do.call(rbind, data_list)

# Save the combined results into the same folder
file_path <- file.path(folder_path, "all_results_us.rds")
saveRDS(all_results_us, file_path)
```

```{r rename_results, echo=FALSE, message=FALSE, warning=FALSE}
# Set the path to your folder
all_results <- all_results_us
```

```{r summary_stats}
# Calculate summary statistics
summary_stats <- all_results %>%
  group_by(n_structures, allocation_type, n_samples) %>%
  summarise(
    mean_ari = mean(recovery_ari),
    sd_ari = sd(recovery_ari),
    prop_perfect = mean(recovery_ari > 0.95),
    prop_failed = mean(recovery_ari < 0.5),
    prop_found_k = mean(found_correct_k),
    mean_structure_recovery = mean(mean_structure_recovery),
    .groups = 'drop'
  )

# Add dimension info for each structure
for (k in n_structures_values) {
  patterns <- allocation_patterns[[paste0("k", k)]]
  for (pattern_name in names(patterns)) {
    dims <- patterns[[pattern_name]]
    for (s in 1:k) {
      col_name <- paste0("struct", s, "_dims")
      summary_stats[summary_stats$n_structures == k & 
                   summary_stats$allocation_type == pattern_name, col_name] <- dims[s]
    }
  }
}

# Display results by k
for (k in n_structures_values) {
  cat(sprintf("\n### Recovery Performance for k = %d structures\n", k))
  
  k_summary <- summary_stats %>%
    filter(n_structures == k) %>%
    arrange(allocation_type, n_samples)
  
  # Create display table
  display_cols <- c("allocation_type", "n_samples", "mean_ari", "prop_found_k")
  
  # Add dimension columns
  for (s in 1:k) {
    display_cols <- c(display_cols, paste0("struct", s, "_dims"))
  }
  
  k_display <- k_summary[, display_cols]
  
  print(knitr::kable(k_display, digits = 3,
                     caption = sprintf("Results for k=%d structures", k)))
}
```

## 4.2 Main Visualisation: Equal vs Unequal Allocation

```{r main_comparison, fig.height=10}
# Create comparison plot
p_comparison <- ggplot(summary_stats, 
                      aes(x = tools::toTitleCase(allocation_type), y = mean_ari, 
                          fill = factor(n_samples))) +
  geom_bar(stat = "identity", position = "dodge", alpha = 0.8) +
  geom_errorbar(aes(ymin = mean_ari - sd_ari, ymax = mean_ari + sd_ari),
                position = position_dodge(width = 0.9), width = 0.25) +
  facet_wrap(~ n_structures, ncol = 2, scales = "free_x", 
             labeller = labeller(n_structures = function(x) paste(x, "structures"))) +
  scale_fill_viridis_d(name = "Sample\nSize", option = "viridis") +
  labs(title = "Recovery Performance: Equal vs Unequal Dimension Allocation",
       subtitle = sprintf("All conditions at %.1fσ separation", separation_factor),
       x = "Allocation Pattern",
       y = "Mean Recovery ARI (± SD)") +
  theme_minimal() +
 theme(plot.title = element_text(hjust = 0.5, size = 12, face = "bold", family = "Times"),
        plot.subtitle = element_text(hjust = 0.5, size = 10, family = "Times"),
        axis.text = element_text(size = 10, family = "Times"),
        axis.title = element_text(size = 12, face = "bold", family = "Times"),
        legend.title = element_text(size = 12, face = "bold", family = "Times"),
        strip.text = element_text(size = 12, family = "Times", face = "bold"),
        legend.text = element_text(size = 10, family = "Times"),
        legend.spacing = unit(1.5, "cm"))

print(p_comparison)

ggsave(paste0("1k_allocationrecovery.pdf"), plot = p_comparison, device = cairo_pdf, width = 6, height = 4.5, units = "in")
```

## 4.3 Structure-Specific Recovery

```{r structure_recovery, fig.height=12}
# Prepare data for structure-specific analysis
struct_recovery_data <- data.frame()

for (k in n_structures_values) {
  k_data <- all_results %>%
    filter(n_structures == k) %>%
    select(n_structures, allocation_type, n_samples, replication, 
           starts_with("struct")) %>%
    pivot_longer(cols = starts_with("struct") & ends_with("_recovery"),
                 names_to = "structure",
                 values_to = "recovery") %>%
    mutate(structure_num = as.numeric(gsub("struct(\\d)_recovery", "\\1", structure)))
  
  # Add dimension info
  for (s in 1:k) {
    dims_col <- paste0("struct", s, "_dims")
    k_data[k_data$structure_num == s, "n_dims"] <- k_data[k_data$structure_num == s, dims_col]
  }
  
  struct_recovery_data <- rbind(struct_recovery_data, k_data)
}

# Calculate summary
struct_recovery_summary <- struct_recovery_data %>%
  group_by(n_structures, allocation_type, n_samples, structure_num, n_dims) %>%
  summarise(mean_recovery = mean(recovery, na.rm = TRUE),
            .groups = 'drop')

# Create plot
p_struct <- ggplot(struct_recovery_summary,
                  aes(x = n_dims, y = mean_recovery, 
                      colour = allocation_type,
                      shape = factor(n_samples))) +
  geom_point(size = 3, alpha = 0.8) +
  geom_line(aes(group = interaction(allocation_type, n_samples)), alpha = 0.5) +
  facet_wrap(~ n_structures, 
             labeller = labeller(n_structures = function(x) paste(x, "structures"))) +
  scale_x_continuous(breaks = c(5, 10, 15, 20, 25, 30, 40, 50, 60, 70, 80)) +
  scale_colour_viridis_d(name = "Allocation\nPattern", option = "plasma") +
  scale_shape_manual(name = "Sample\nSize", values = c(16, 17, 18)) +
  labs(title = "Structure Recovery by Number of Dimensions",
       subtitle = "How well is each structure recovered based on its dimension allocation?",
       x = "Number of Dimensions Allocated to Structure",
       y = "Mean Structure Recovery Rate") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
        plot.subtitle = element_text(hjust = 0.5, size = 12))

print(p_struct)
```

## 4.4 Extreme Allocations Analysis

```{r extreme_analysis}
# Focus on extreme allocations
extreme_data <- summary_stats %>%
  filter(allocation_type == "extreme") %>%
  arrange(n_structures, n_samples)

cat("\n### Performance with Extreme Allocations\n")
cat("These represent the most unbalanced dimension distributions:\n\n")

for (k in n_structures_values) {
  k_extreme <- extreme_data %>% filter(n_structures == k)
  if (nrow(k_extreme) > 0) {
    cat(sprintf("k=%d: ", k))
    dims <- allocation_patterns[[paste0("k", k)]][["extreme"]]
    cat(sprintf("Dimensions = (%s)\n", paste(dims, collapse=", ")))
    cat(sprintf("  Mean ARI across all n: %.3f\n", mean(k_extreme$mean_ari)))
    cat(sprintf("  Ever finds correct k: %s\n", 
                ifelse(any(k_extreme$prop_found_k > 0), "Yes", "No")))
  }
}

# Compare extreme vs equal
comparison_data <- summary_stats %>%
  filter(allocation_type %in% c("equal", "extreme")) %>%
  group_by(n_structures, allocation_type) %>%
  summarise(mean_ari = mean(mean_ari),
            .groups = 'drop') %>%
  pivot_wider(names_from = allocation_type, values_from = mean_ari) %>%
  mutate(performance_drop = equal - extreme,
         percent_drop = 100 * performance_drop / equal)

cat("\n### Performance Drop from Equal to Extreme Allocation:\n")
print(comparison_data)
```

## 4.5 Minimum Viable Dimensions

```{r min_dims_analysis}
# Analyze minimum dimensions needed for detection
dim_threshold_data <- struct_recovery_data %>%
  group_by(n_dims, n_samples) %>%
  summarise(
    mean_recovery = mean(recovery, na.rm = TRUE),
    prop_detected = mean(recovery > 0.8, na.rm = TRUE),
    .groups = 'drop'
  ) %>%
  filter(!is.na(n_dims))

# Plot detection threshold
p_threshold <- ggplot(dim_threshold_data,
                     aes(x = n_dims, y = prop_detected, 
                         colour = factor(n_samples))) +
  geom_line(size = 1.2) +
  geom_point(size = 3) +
  geom_hline(yintercept = 0.5, linetype = "dashed", alpha = 0.5) +
  geom_vline(xintercept = c(10, 20, 30), linetype = "dotted", alpha = 0.3) +
  scale_colour_viridis_d(name = "Sample\nSize", option = "viridis") +
  labs(title = "Structure Detection Probability by Dimension Count",
       subtitle = "What's the minimum number of dimensions needed for reliable detection?",
       x = "Number of Dimensions",
       y = "Proportion Successfully Detected (>80% recovery)") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
        plot.subtitle = element_text(hjust = 0.5, size = 12))

print(p_threshold)

# Find thresholds
thresholds <- dim_threshold_data %>%
  filter(prop_detected >= 0.5) %>%
  group_by(n_samples) %>%
  summarise(min_dims = min(n_dims),
            .groups = 'drop')

cat("\n### Minimum Dimensions for 50% Detection Success:\n")
print(thresholds)
```
