---
title: "Simulation Study: Effect of Dimensionality on Clustering Structure Recovery"
author: "Your Name"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, 
                      fig.width = 12, fig.height = 8, cache = TRUE)
```

# Executive Summary

This simulation study investigates how the **number of dimensions (p)** affects the ability to recover clustering structures. We maintain a design where half the dimensions contain one clustering structure and half contain another, but vary the total number of dimensions from 6 to 100.

**Key Questions**:
1. How does the curse of dimensionality affect structure recovery?
2. Does the critical separation threshold depend on the number of dimensions?
3. What is the relationship between sample size, dimensionality, and recovery performance?
4. Does higher dimensionality help (more evidence) or hurt (more noise)?

# 1. Methods

## 1.1 Experimental Design

We maintain the two-structure design, but vary the number of dimensions:
- **Structure 1**: 3 clusters in the first p/2 dimensions
- **Structure 2**: 4 clusters in the last p/2 dimensions

### Experimental Grid:
- **Dimensions (p)**: 6, 10, 20, 30, 50, 100
- **Sample sizes**: 100, 250, 500, 1000, 2000
- **Separation factors**: 3.0σ, 2.0σ, 1.5σ, 1.0σ
- **Replications**: Easily adjustable (see parameter section)

# 2. Implementation

```{r libraries}
# Load required libraries
library(mvtnorm)      # For multivariate normal data generation
library(cluster)      # For PAM clustering
library(mclust)       # For model-based clustering
library(ggplot2)      # For visualisation
library(dplyr)        # For data manipulation
library(tidyr)        # For data reshaping
library(viridis)      # For colour scales
library(gridExtra)    # For arranging plots
library(scales)       # For formatting
library(foreach)      # For parallel loops
library(doParallel)   # For parallel backend
library(iterators)    # For iteration support

# Set seed for reproducibility
set.seed(42)
```

```{r key_parameter}
# Set to 1 for test run, 25 for full simulation
n_replications <- 25

# Set parallel processing preferences
use_parallel <- TRUE  # Set to FALSE to run sequentially
n_cores_to_use <- detectCores() - 2  # Leave 2 cores free for system

cat("SIMULATION CONFIGURATION:\n")
cat("Number of replications per condition:", n_replications, "\n")
cat("Parallel processing:", ifelse(use_parallel, "ENABLED", "DISABLED"), "\n")
if (use_parallel) {
  cat("Cores available:", detectCores(), "\n")
  cat("Cores to use:", n_cores_to_use, "\n")
}

if (n_replications == 1) {
  cat("\n*** Running in TEST MODE with 1 replication per condition ***\n")
  cat("*** Change n_replications to 25 for full simulation ***\n")
} else {
  cat("\nRunning full simulation with", n_replications, "replications per condition\n")
}
```

```{r cache_setup}
# Set up caching directory with replication info
cache_base_dir <- "dimensionality_simulation_cache"
cache_dir <- file.path(cache_base_dir, paste0("nrep_", n_replications))

if (!dir.exists(cache_base_dir)) {
  dir.create(cache_base_dir)
}
if (!dir.exists(cache_dir)) {
  dir.create(cache_dir)
  cat("\nCreated cache directory:", cache_dir, "\n")
} else {
  cat("\nUsing existing cache directory:", cache_dir, "\n")
}

# Function to generate cache filename for a specific condition
get_cache_filename <- function(p_dims, n_samples, separation_factor) {
  file.path(cache_dir, sprintf("results_p%d_n%d_sep%.1f.rds", p_dims, n_samples, separation_factor))
}

# Function to check if results exist for a specific condition
check_condition_cache <- function(p_dims, n_samples, separation_factor) {
  cache_file <- get_cache_filename(p_dims, n_samples, separation_factor)
  if (file.exists(cache_file)) {
    # Return TRUE only if file size > 0 bytes (not empty)
    return(file.info(cache_file)$size > 0)
  } else {
    return(FALSE)
  }
}

# Function to save results for a specific condition
save_condition_results <- function(results, p_dims, n_samples, separation_factor) {
  cache_file <- get_cache_filename(p_dims, n_samples, separation_factor)
  saveRDS(results, cache_file)
  cat("Saved results to cache:", cache_file, "\n")
  return(results)
}

# Function to load results for a specific condition
load_condition_results <- function(p_dims, n_samples, separation_factor) {
  cache_file <- get_cache_filename(p_dims, n_samples, separation_factor)
  if (file.exists(cache_file)) {
    return(readRDS(cache_file))
  } else {
    return(NULL)
  }
}
```

```{r parameters}
# Experimental grid
p_values <- c(10, 20, 30, 50, 100)
sample_sizes <- c(100, 250, 500, 1000, 2000)
separation_factors <- c(3.0, 2.0, 1.0)

# Fixed parameters
cluster_variance <- 0.3

# Calculate total number of conditions
n_conditions <- length(p_values) * length(sample_sizes) * length(separation_factors)
total_runs <- n_conditions * n_replications

# Create condition grid
condition_grid <- expand.grid(
  p_dims = p_values,
  n_samples = sample_sizes,
  separation_factor = separation_factors,
  stringsAsFactors = FALSE
)

# Check which conditions are already cached
condition_grid$is_cached <- mapply(check_condition_cache,
                                   condition_grid$p_dims,
                                   condition_grid$n_samples,
                                   condition_grid$separation_factor)

n_cached <- sum(condition_grid$is_cached)
n_to_compute <- sum(!condition_grid$is_cached)

cat("\nExperimental design:\n")
cat("- Dimensions (p):", paste(p_values, collapse=", "), "\n")
cat("- Sample sizes:", paste(sample_sizes, collapse=", "), "\n")
cat("- Separation factors:", paste(separation_factors, collapse=", "), "σ\n")
cat("- Total conditions:", n_conditions, "\n")
cat("- Conditions already cached:", n_cached, "\n")
cat("- Conditions to compute:", n_to_compute, "\n")
cat("- Replications per condition:", n_replications, "\n")
cat("- Total simulation runs needed:", n_to_compute * n_replications, "\n")
```

```{r data_generation_functions}
# Modified functions to handle variable p

generate_cluster_means_p <- function(separation_factor, cluster_variance, p_dims) {
  base_separation <- separation_factor * sqrt(cluster_variance)
  
  # Structure 1 (first half of dimensions) - 3 clusters
  p_half <- as.integer(p_dims / 2)
  p1 <- p_half
  means_1 <- list(
    cluster_1 = runif(p1, -1.5, -1) * base_separation,
    cluster_2 = runif(p1, 1, 1.5) * base_separation,
    cluster_3 = rep(0, p1)
  )
  
  # Structure 2 (second half of dimensions) - 4 clusters 
  p_half <- as.integer(p_dims / 2)
  p2 <- p_half
  base_vec <- runif(p2, -0.4, 0.4) * base_separation
  means_2 <- list(
    cluster_1 = runif(p2, -1.4, -1) * base_separation,
    cluster_2 = runif(p2, 1, 1.4) * base_separation,
    cluster_3 = base_vec,
    cluster_4 = -base_vec
  )

  
  return(list(means_1 = means_1, means_2 = means_2))
}

generate_dataset_p <- function(separation_factor, cluster_variance, n_samples, p_dims) {
  
  # Fixed cluster proportions
  cluster_props_1 <- c(0.4, 0.35, 0.25)
  cluster_props_2 <- c(0.3, 0.25, 0.25, 0.2)
  
  # Generate cluster assignments
  cluster_1_assignments <- sample(1:3, n_samples, replace = TRUE, prob = cluster_props_1)
  cluster_2_assignments <- sample(1:4, n_samples, replace = TRUE, prob = cluster_props_2)
  
  # Get cluster means
  means_list <- generate_cluster_means_p(separation_factor, cluster_variance, p_dims)
  means_1 <- means_list$means_1
  means_2 <- means_list$means_2
  
  # Define covariance matrices
  p_half <- as.integer(p_dims / 2)
  p1 <- p_half
  p2 <- p_half
  
  covs_1 <- list(
    cluster_1 = diag(rep(cluster_variance, p1)),
    cluster_2 = diag(rep(cluster_variance, p1)),
    cluster_3 = diag(rep(cluster_variance, p1))
  )
  
  covs_2 <- list(
    cluster_1 = diag(rep(cluster_variance, p2)),
    cluster_2 = diag(rep(cluster_variance, p2)),
    cluster_3 = diag(rep(cluster_variance, p2)),
    cluster_4 = diag(rep(cluster_variance, p2))
  )
  
  # Initialise data matrix
  data_matrix <- matrix(NA, nrow = n_samples, ncol = p_dims)
  
  # Generate data for Structure 1
  for (i in 1:n_samples) {
    cluster_id <- cluster_1_assignments[i]
    data_matrix[i, 1:p1] <- rmvnorm(1, mean = means_1[[cluster_id]], 
                                    sigma = covs_1[[cluster_id]])
  }
  
  # Generate data for Structure 2
  for (i in 1:n_samples) {
    cluster_id <- cluster_2_assignments[i]
    data_matrix[i, (p1+1):p_dims] <- rmvnorm(1, mean = means_2[[cluster_id]], 
                                             sigma = covs_2[[cluster_id]])
  }
  
  return(data_matrix)
}
```

```{r analysis_functions}
# Modified analysis function to handle variable p

analyze_dataset_p <- function(data_matrix, p_dims) {
  n_dims <- ncol(data_matrix)
  
  # Step 1: MCLUST on each dimension
  mclust_results <- list()
  cluster_assignments <- matrix(NA, nrow = nrow(data_matrix), ncol = n_dims)
  
  for(i in 1:n_dims) {
    mclust_fit <- Mclust(data_matrix[, i], G = 1:9, verbose = FALSE)
    mclust_results[[i]] <- mclust_fit
    cluster_assignments[, i] <- mclust_fit$classification
  }
  
  # Step 2: Compute ARI matrix
  ari_matrix <- matrix(NA, nrow = n_dims, ncol = n_dims)
  for(i in 1:n_dims) {
    for(j in 1:n_dims) {
      if(i == j) {
        ari_matrix[i, j] <- 1.0
      } else {
        ari_matrix[i, j] <- adjustedRandIndex(cluster_assignments[, i], 
                                              cluster_assignments[, j])
      }
    }
  }
  
  # Step 3: Hierarchical clustering
  dissim_matrix <- 1 - ari_matrix
  diag(dissim_matrix) <- 0
  dissim_dist <- as.dist(dissim_matrix)
  
  p_half <- as.integer(p_dims / 2)
  max_k <- min(6, floor(p_half))

  index <- "silhouette"
  
  k_values <- 2:6
  index_values <- numeric(length(k_values))
  hclust_results <- vector("list", length(k_values))
  
  for (i in seq_along(k_values)) {
    k <- k_values[i]
    hclust_fit <- hclust(dissim_dist, method = "single")
    clusters <- cutree(hclust_fit, k = k)
    hclust_results[[i]] <- clusters
    sil <- silhouette(clusters, dist = dissim_dist)
    index_values[i] <- mean(sil[, "sil_width"])
    cat(sprintf("\nk = %d: Average silhouette width = %.3f", k, index_values[i]))
  }
  
  optimal_idx <- which.max(index_values)
  optimal_k <- k_values[optimal_idx]
  optimal_hclust <- hclust_results[[optimal_idx]]
  
  cat("\nOptimal number of dimension groups (silhouette):", optimal_k, "\n")
  
  # Step 4: Calculate recovery ARI
  # True structure labels
  p_half <- as.integer(as.integer(p_dims / 2))
  true_structure_labels <- c(rep(1, p_half), rep(2, p_half))
  recovery_ari <- adjustedRandIndex(optimal_hclust, true_structure_labels)
  
  # Additional metrics
  detected_clusters <- sapply(mclust_results, function(x) x$G)
  
  return(list(
    recovery_ari = recovery_ari,
    optimal_k = optimal_k,
    index = index,
    index_value = as.numeric(index_values[optimal_idx]),
    detected_clusters = detected_clusters,
    hierarchical_clustering = optimal_hclust,
    dims1_correct = sum(detected_clusters[1:(p_half)] == 3),
    dims2_correct = sum(detected_clusters[(p_half+1):p_dims] == 4)
  ))
}
```

```{r worker_function}
# Function to run a single condition (for parallel processing)
run_single_condition <- function(condition_row, n_replications, cluster_variance, cache_dir) {
  
  p_dims <- condition_row$p_dims
  n_samples <- condition_row$n_samples
  sep_factor <- condition_row$separation_factor
  
  # Check cache first
  if (check_condition_cache(p_dims, n_samples, sep_factor)) {
    return(NULL)  # Signal that this was cached
  }
  
  # Run the simulation for this condition
  condition_results <- data.frame()
  
  for (rep in 1:n_replications) {
    # Generate and analyse dataset
    data_matrix <- generate_dataset_p(sep_factor, cluster_variance, n_samples, p_dims)
    analysis_result <- analyze_dataset_p(data_matrix, p_dims)
    
    p_half <- as.integer(p_dims / 2)
    
    # Store results
    condition_results <- rbind(condition_results, data.frame(
      p_dims = p_dims,
      n_samples = n_samples,
      separation_factor = sep_factor,
      replication = rep,
      recovery_ari = analysis_result$recovery_ari,
      optimal_k = analysis_result$optimal_k,
      index = analysis_result$index,
      index_value = analysis_result$index_value,
      dims1_correct = analysis_result$dims1_correct,
      dims2_correct = analysis_result$dims2_correct,
      prop_dims1_correct = analysis_result$dims1_correct / (p_half),
      prop_dims2_correct = analysis_result$dims2_correct / (p_half)
    ))
  }
  
  cat(sprintf("Completed condition: p=%d, n=%d, sep=%.1f, reps=%d\n", p_dims, n_samples, sep_factor, n_replications))

  
  # Cache results
  save_condition_results(condition_results, p_dims, n_samples, sep_factor)
  
  
  return(condition_results)
}
```

# 3. Run Simulation

```{r run_simulation, cache=FALSE}
# Main simulation execution with parallel support and caching
library(foreach)
library(doParallel)

start_time <- Sys.time()
cat("\nStarting dimensionality × sample size × separation simulation...\n")
cat("Time started:", format(start_time, "%Y-%m-%d %H:%M:%S"), "\n\n")

# Identify conditions not yet computed/cached
conditions_to_compute <- condition_grid[!condition_grid$is_cached, ]

all_results_file <- file.path(cache_dir, "all_results_d.rds")

if (nrow(conditions_to_compute) == 0) {
  cat("All conditions are already cached! Loading combined results...\n")
  all_results <- readRDS(all_results_file)
  
} else if (!use_parallel || nrow(conditions_to_compute) < 4) {
  # Sequential processing if parallel is disabled or few conditions to compute
  cat("Running simulation in SEQUENTIAL mode...\n")
  all_results <- data.frame()
  
  for (i in seq_len(nrow(conditions_to_compute))) {
    cond <- conditions_to_compute[i, ]
    cat(sprintf("Computing condition %d/%d: p=%d, n=%d, sep=%.1fσ\n",
                i, nrow(conditions_to_compute), cond$p_dims, cond$n_samples, cond$separation_factor))
    
    res <- run_single_condition(cond, n_replications, cluster_variance, cache_dir)
    if (!is.null(res)) {
      all_results <- rbind(all_results, res)
    }
  }
  
} else {
  # Parallel processing setup
  cat(sprintf("Running simulation in PARALLEL mode with %d cores...\n", n_cores_to_use))
  cat(sprintf("Processing %d conditions in parallel\n", nrow(conditions_to_compute)))
  
  cl <- parallel::makeCluster(n_cores_to_use)
  doParallel::registerDoParallel(cl)
  
  # Export necessary objects and functions to workers
  parallel_exports <- c(
    "generate_cluster_means_p", "generate_dataset_p", "analyze_dataset_p",
    "check_condition_cache", "save_condition_results", "get_cache_filename",
    "cluster_variance", "n_replications", "cache_dir", "run_single_condition"
  )
  parallel::clusterExport(cl, parallel_exports, envir = environment())
  
  # Load required packages on workers
  parallel::clusterEvalQ(cl, {
    library(mvtnorm)
    library(cluster)
    library(mclust)
    library(dplyr)
    library(parallel)
    library(foreach)
    library(doParallel)
  })
  
  # Progress bar setup
  pb <- txtProgressBar(min = 0, max = nrow(conditions_to_compute), style = 3)
  progress <- function(n) setTxtProgressBar(pb, n)
  opts <- list(progress = progress)
  
  # Run simulation in parallel
  parallel_results <- foreach(
  i = seq_len(nrow(conditions_to_compute)),
  .combine = 'rbind',
  .options.snow = opts,
  .errorhandling = 'pass'  # keep this for now
) %do% {

  cond <- conditions_to_compute[i, ]
  tryCatch({
    cat(sprintf(">>> Starting: p = %d, n = %d, sep = %.1f\n",
                cond$p_dims, cond$n_samples, cond$separation_factor))
    
    result <- run_single_condition(cond, n_replications, cluster_variance, cache_dir)

    cat(sprintf("✓ Finished: p = %d, n = %d, sep = %.1f\n",
                cond$p_dims, cond$n_samples, cond$separation_factor))
    
    result
  }, error = function(e) {
    cat(sprintf("⚠️ Error in p = %d, n = %d, sep = %.1f: %s\n",
                cond$p_dims, cond$n_samples, cond$separation_factor, e$message))
    NULL
  })
}

  
  close(pb)
  parallel::stopCluster(cl)
  
  # Combine results, filtering out NULLs from cached conditions
  if (!is.null(parallel_results)) {
    all_results <- do.call(rbind, parallel_results[!sapply(parallel_results, is.null)])
  } else {
    all_results <- data.frame()
  }
}

# Load all cached results including new ones to ensure completeness
cat("\nLoading all results from cache...\n")
complete_results <- data.frame()

for (i in seq_len(nrow(condition_grid))) {
  cond <- condition_grid[i, ]
  cached_data <- load_condition_results(cond$p_dims, cond$n_samples, cond$separation_factor)
  
  if (!is.null(cached_data)) {
    complete_results <- rbind(complete_results, cached_data)
  }
}

all_results <- complete_results

end_time <- Sys.time()
total_duration <- difftime(end_time, start_time, units = "mins")

# Save combined results
saveRDS(all_results, all_results_file)
cat("Combined results saved to:", all_results_file, "\n")

cat("\n=== SIMULATION COMPLETE ===\n")
cat("Total time for this run:", sprintf("%.1f minutes\n", total_duration))
cat("Total results in dataset:", nrow(all_results), "\n")
cat("Results cached in:", cache_dir, "\n")

# Final cache status report
final_cached <- sum(condition_grid$is_cached)
cat(sprintf("Cache status: %d/%d conditions cached\n", final_cached, nrow(condition_grid)))
```

# 4. Results

## 4.1 Summary Statistics

```{r join-results, echo=FALSE, message=FALSE, warning=FALSE}
# Set the path to your folder
folder_path <- "dimensionality_simulation_cache/nrep_25"

# List all RDS files in the folder
file_list <- list.files(path = folder_path, pattern = "^results.*\\.rds$", full.names = TRUE)

# Read all RDS files and store them in a list
data_list <- lapply(file_list, readRDS)

# Assuming each RDS contains a data frame, combine them
all_results_d <- do.call(rbind, data_list)

# Save the combined results into the same folder
file_path <- file.path(folder_path, "all_results_d.rds")
saveRDS(all_results_d, file_path)
```

```{r summary_stats}
library(tidyverse)
# convert to all results
all_results <- all_results_d

# Calculate summary statistics
summary_stats <- all_results %>%
  group_by(p_dims, n_samples, separation_factor) %>%
  summarise(
    mean_ari = mean(recovery_ari),
    sd_ari = sd(recovery_ari),
    prop_perfect = mean(recovery_ari == 1),
    prop_failed = mean(recovery_ari < 0.5),
    mean_silhouette = mean(index_value),
    mean_optimal_k = mean(optimal_k),
    prop_k_correct = mean(optimal_k == 2),
    .groups = 'drop'
  ) %>%
  mutate(n_p_ratio = n_samples / p_dims)

# Create summary table for each p value
for (p in p_values) {
  cat(sprintf("\n### Recovery Performance for p = %d dimensions\n", p))
  
  p_summary <- summary_stats %>%
    filter(p_dims == p) %>%
    select(n_samples, separation_factor, mean_ari) %>%
    pivot_wider(names_from = separation_factor, 
                values_from = mean_ari,
                names_prefix = "Sep_")
  print(p_summary)
  print(knitr::kable(p_summary, digits = 3,
                     caption = sprintf("Mean Recovery ARI for p = %d", p),
                     col.names = c("Sample Size", "3.0σ", "2.0σ", "1.0σ")))
}
```

## 4.2 Main Visualisation: The p × n × separation Landscape

```{r main_heatmap, fig.height=10}
# Create faceted heatmap
p_heatmap <- ggplot(summary_stats, aes(x = factor(n_samples), 
                                       y = factor(separation_factor), 
                                       fill = mean_ari)) +
  geom_tile(colour = "white", size = 0.5) +
  geom_text(aes(label = sprintf("%.3f", mean_ari)), 
            colour = ifelse(summary_stats$mean_ari > 0.5, "white", "black"),
            size = 3, family = "Times") +
  facet_wrap(~ p_dims, ncol = 2, 
             labeller = labeller(p_dims = function(x) paste("p =", x, "dimensions"))) +
  scale_fill_viridis(name = "Mean\nRecovery\nARI", 
                     limits = c(0, 1),
                     option = "plasma") +
  scale_y_discrete(labels = function(x) paste0(x, "σ")) +
  scale_x_discrete(labels = function(x) format(as.numeric(x), big.mark = ",")) +
  labs(title = "Structure Recovery Performance: The Curse of Dimensionality",
       subtitle = paste("Based on", n_replications, 
                       ifelse(n_replications == 1, "replication", "replications"), 
                       "per condition"),
       x = "Sample Size",
       y = "Separation Factor") +
  theme_minimal() +
 theme(plot.title = element_text(hjust = 0.5, size = 12, face = "bold", family = "Times"),
        plot.subtitle = element_text(hjust = 0.5, size = 10, family = "Times"),
        axis.text = element_text(size = 10, family = "Times"),
        axis.title = element_text(size = 12, face = "bold", family = "Times"),
        legend.title = element_text(size = 12, face = "bold", family = "Times"),
        legend.text = element_text(size = 10, family = "Times"),
       strip.text = element_text(size = 12, family = "Times", face = "bold"),
        panel.grid = element_blank(),
        panel.spacing = unit(1, "lines"))

ggsave(paste0("1i_meanARIheatmap.pdf"), plot = p_heatmap, device = cairo_pdf, width = 6, height = 8, units = "in")
```

## 4.3 The n/p Ratio Analysis

```{r np_ratio_plot}
# Analyse performance by n/p ratio
p_ratio <- ggplot(summary_stats, aes(x = n_p_ratio, y = mean_ari, 
                                     colour = factor(separation_factor))) +
  geom_line(aes(group = interaction(p_dims, separation_factor)), 
            alpha = 0.3, size = 0.8) +
  geom_point(aes(shape = factor(p_dims)), size = 3, alpha = 0.8) +
  scale_x_log10(breaks = c(1, 2, 5, 10, 20, 50, 100, 200),
                labels = c(1, 2, 5, 10, 20, 50, 100, 200)) +
  scale_colour_viridis_d(name = "Separation", 
                        labels = function(x) paste0(x, "σ"),
                        option = "plasma") +
  scale_shape_manual(name = "Dimensions\n(p)", 
                     values = c(16, 17, 15, 18, 8, 4)) +
  geom_vline(xintercept = c(10, 20), linetype = "dashed", alpha = 0.5) +
  geom_hline(yintercept = 0.8, linetype = "dashed", alpha = 0.5, colour = "darkgreen") +
  labs(title = "Performance by Sample Size to Dimension Ratio (n/p)",
       subtitle = "Vertical lines at n/p = 10 and 20; horizontal line at ARI = 0.8",
       x = "n/p Ratio (log scale)",
       y = "Mean Recovery ARI") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
        plot.subtitle = element_text(hjust = 0.5, size = 12),
        legend.position = "right")

print(p_ratio)
```

## 4.4 Dimensionality Effect at Fixed Sample Sizes

```{r dim_effect_plot}
# Show how performance changes with p at fixed n
p_effect <- ggplot(summary_stats, aes(x = p_dims, y = mean_ari, 
                                      colour = factor(n_samples))) +
  geom_line(size = 1.2) +
  geom_point(size = 3) +
  facet_wrap(~ separation_factor, ncol = 2,
             labeller = labeller(separation_factor = function(x) paste(x, "σ separation"))) +
  scale_x_log10(breaks = p_values, labels = p_values) +
  scale_colour_viridis_d(name = "Sample\nSize", option = "viridis") +
  geom_hline(yintercept = 0.8, linetype = "dashed", alpha = 0.5) +
  labs(title = "Effect of Dimensionality on Recovery Performance",
       subtitle = "How does increasing p affect recovery at fixed sample sizes?",
       x = "Number of Dimensions (p) - log scale",
       y = "Mean Recovery ARI") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
        plot.subtitle = element_text(hjust = 0.5, size = 12),
        strip.text = element_text(size = 11, face = "bold"),
        legend.position = "right")

print(p_effect)
```

## 4.5 Critical Thresholds Table

```{r threshold_table}
# Find critical n/p ratios for reliable recovery
threshold_analysis <- summary_stats %>%
  filter(mean_ari >= 0.8) %>%
  group_by(p_dims) %>%
  summarise(
    min_n_for_sep3 = min(n_samples[separation_factor == 3.0], na.rm = TRUE),
    min_n_for_sep2 = min(n_samples[separation_factor == 2.0], na.rm = TRUE),
    min_n_for_sep1 = min(n_samples[separation_factor == 1.0], na.rm = TRUE),
    .groups = 'drop'
  ) %>%
  mutate(
    ratio_sep3 = ifelse(is.finite(min_n_for_sep3), min_n_for_sep3 / p_dims, NA),
    ratio_sep2 = ifelse(is.finite(min_n_for_sep2), min_n_for_sep2 / p_dims, NA),
    ratio_sep1 = ifelse(is.finite(min_n_for_sep1), min_n_for_sep1 / p_dims, NA)
  )

# Format for display
threshold_display <- threshold_analysis %>%
  mutate(across(starts_with("min_n"), ~ifelse(is.finite(.), as.character(.), ">2000"))) %>%
  select(p_dims, min_n_for_sep3, min_n_for_sep2, min_n_for_sep1)

knitr::kable(threshold_display,
             caption = "Minimum Sample Size Required for 80% Recovery Success",
             col.names = c("Dimensions (p)", "At 3.0σ", "At 2.0σ", "At 1.σ"))

# n/p ratio table
ratio_display <- threshold_analysis %>%
  select(p_dims, ratio_sep3, ratio_sep2, ratio_sep1) %>%
  mutate(across(starts_with("ratio"), ~ifelse(is.na(.), "-", sprintf("%.3f", .))))

knitr::kable(ratio_display,
             caption = "Minimum n/p Ratio Required for 80% Recovery Success",
             col.names = c("Dimensions (p)", "At 3.0σ", "At 2.0σ", "At 1.0σ"))
```
