---
title: "Discovering Multiple Clustering Structures: A Step-by-Step Analysis"
author: "Your Name"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, 
                      fig.width = 12, fig.height = 8, cache = TRUE)
```

# Introduction

This document demonstrates a novel approach for detecting and separating multiple independent clustering structures within high-dimensional data. We walk through a complete analysis of a single synthetic dataset designed to have two distinct clustering patterns in different subsets of dimensions.

## The Challenge

Traditional clustering algorithms assume a single clustering structure exists across all dimensions. However, real-world data often contains multiple, independent groupings. For instance:
- Customer data might cluster by purchasing behaviour in some variables and by demographics in others
- Gene expression data might show different clustering patterns for different biological pathways
- Sensor data might have spatial patterns in some dimensions and temporal patterns in others

## Our Approach

We demonstrate a methodology that:
1. Applies clustering independently to each dimension
2. Measures similarity between these dimension-wise clusterings
3. Groups dimensions that share similar clustering patterns
4. Validates whether distinct clustering structures have been successfully identified

## Dataset Design

Our synthetic dataset contains:
- **10 dimensions** total
- **Structure 1**: 3 clusters present in dimensions 1-5
- **Structure 2**: 4 clusters present in dimensions 6-10
- **High separation** (6σ) to clearly demonstrate the methodology

This creates a challenging scenario where traditional clustering must choose which structure to optimise for, as both cannot be captured simultaneously.

# 1. Setup and Libraries

```{r libraries}
# Load required libraries
library(mvtnorm)      # For multivariate normal data generation
library(ggplot2)      # For visualisation
library(cluster)      # For PAM clustering
library(factoextra)   # For clustering visualisation
library(mclust)       # For model-based clustering
library(pheatmap)     # For heatmaps
library(RColorBrewer) # For colour palettes
library(dplyr)        # For data manipulation
library(viridis)      # For colour scales

# Set seed for reproducibility
set.seed(42)
```

# 2. Data Generation

We create synthetic data where two independent clustering structures coexist in different dimensional subspaces.

```{r parameters}
# Dataset parameters
n_samples <- 50               # Number of observations (small for clear visualisation)
separation_factor <- 6.0      # High separation for clear demonstration
cluster_variance <- 0.3       # Within-cluster variance

# Clustering structures
n_clusters_1 <- 3             # Number of clusters in Structure 1 (dims 1-5)
n_clusters_2 <- 4             # Number of clusters in Structure 2 (dims 6-10)
cluster_props_1 <- c(0.4, 0.35, 0.25)      # Cluster proportions for Structure 1
cluster_props_2 <- c(0.3, 0.25, 0.25, 0.2) # Cluster proportions for Structure 2

# Dimension names
dim_names <- paste0("Dim", 1:10)

cat("Dataset Configuration:\n")
cat("- Total dimensions:", 10, "\n")
cat("- Structure 1: 3 clusters in dimensions 1-5\n")
cat("- Structure 2: 4 clusters in dimensions 6-10\n")
cat("- Separation factor:", separation_factor, "standard deviations\n")
cat("- Sample size:", n_samples, "observations\n")
```

```{r generation_functions}
#' Generate cluster means based on separation factor
#' 
#' @param separation_factor Multiplier for cluster separation (higher = more separated)
#' @param cluster_variance Within-cluster variance
#' @return List containing means for both structures
generate_cluster_means <- function(separation_factor, cluster_variance) {
  
  # Calculate base separation distance
  base_separation <- separation_factor * sqrt(cluster_variance)
  
  # Structure 1 means (3 clusters, dims 1-5)
  # Each cluster has a distinct pattern across the 5 dimensions
  means_1 <- list(
    cluster_1 = c(-1, -1.3, -1.2, -1.5, -1.1) * base_separation,
    cluster_2 = c(1, 1.3, 1.2, 1.5, 1.1) * base_separation,
    cluster_3 = c(0, 0, 0, 0, 0)  # Centred at origin
  )
  
  # Structure 2 means (4 clusters, dims 6-10)  
  # Different pattern to ensure independence from Structure 1
  means_2 <- list(
    cluster_1 = c(-1.3, -1, -1.25, -1.4, -1.2) * base_separation,
    cluster_2 = c(1.3, 1, 1.25, 1.4, 1.2) * base_separation,
    cluster_3 = c(-0.3, 0.3, -0.4, 0.4, -0.25) * base_separation,
    cluster_4 = c(0.3, -0.3, 0.4, -0.4, 0.25) * base_separation
  )
  
  return(list(means_1 = means_1, means_2 = means_2))
}

#' Generate the complete dataset with two clustering structures
#' 
#' @param separation_factor Controls cluster separation
#' @param cluster_variance Within-cluster variance
#' @param n_samples Number of observations to generate
#' @param cluster_props_1 Cluster proportions for Structure 1
#' @param cluster_props_2 Cluster proportions for Structure 2
#' @return List containing data matrix and true labels
generate_dataset <- function(separation_factor, cluster_variance, n_samples, 
                           cluster_props_1, cluster_props_2) {
  
  # Generate cluster assignments for each structure independently
  cluster_1_assignments <- sample(1:n_clusters_1, n_samples, 
                                 replace = TRUE, prob = cluster_props_1)
  cluster_2_assignments <- sample(1:n_clusters_2, n_samples, 
                                 replace = TRUE, prob = cluster_props_2)
  
  # Get cluster means for this separation level
  means_list <- generate_cluster_means(separation_factor, cluster_variance)
  means_1 <- means_list$means_1
  means_2 <- means_list$means_2
  
  # Define covariance matrices (spherical clusters)
  covs_1 <- lapply(1:3, function(i) diag(rep(cluster_variance, 5)))
  covs_2 <- lapply(1:4, function(i) diag(rep(cluster_variance, 5)))
  
  # Initialise data matrix
  data_matrix <- matrix(NA, nrow = n_samples, ncol = 10)
  
  # Generate data for Structure 1 (dimensions 1-5)
  for (i in 1:n_samples) {
    cluster_id <- cluster_1_assignments[i]
    data_matrix[i, 1:5] <- rmvnorm(1, 
                                  mean = means_1[[cluster_id]], 
                                  sigma = covs_1[[cluster_id]])
  }
  
  # Generate data for Structure 2 (dimensions 6-10)
  for (i in 1:n_samples) {
    cluster_id <- cluster_2_assignments[i]
    data_matrix[i, 6:10] <- rmvnorm(1, 
                                   mean = means_2[[cluster_id]], 
                                   sigma = covs_2[[cluster_id]])
  }
  
  # Return data and true labels
  return(list(
    data = data_matrix,
    true_labels_1 = cluster_1_assignments,
    true_labels_2 = cluster_2_assignments
  ))
}
```

```{r generate_data}
# Generate the dataset
cat("\nGenerating dataset with two independent clustering structures...\n")
dataset <- generate_dataset(
  separation_factor = separation_factor, 
  cluster_variance = cluster_variance, 
  n_samples = n_samples,
  cluster_props_1 = cluster_props_1, 
  cluster_props_2 = cluster_props_2
)

# Extract data and labels
data_matrix <- dataset$data
colnames(data_matrix) <- dim_names

cat("Dataset created successfully.\n")
cat("- Dimensions:", ncol(data_matrix), "\n")
cat("- Samples:", nrow(data_matrix), "\n")
```

# 3. Visualise the Challenge: Two Hidden Structures

This visualisation demonstrates the fundamental challenge: when data is ordered by one clustering structure, the other appears completely scrambled. This illustrates why traditional clustering methods fail to capture both structures simultaneously.

```{r visualise_structures, fig.height=10}
# Create annotation data frames for both structures
annotation_both <- data.frame(
  Structure1 = factor(dataset$true_labels_1, levels = 1:3),
  Structure2 = factor(dataset$true_labels_2, levels = 1:4)
)

# Define consistent colours for annotations
ann_colors <- list(
  Structure1 = c("1" = "#E41A1C", "2" = "#377EB8", "3" = "#4DAF4A"),
  Structure2 = c("1" = "#984EA3", "2" = "#FF7F00", "3" = "#FFFF33", "4" = "#A65628")
)

# Order rows by Structure 1 clustering
order_struct1 <- order(dataset$true_labels_1)
data_ordered1 <- data_matrix[order_struct1, ]
annotation_ordered1 <- annotation_both[order_struct1, , drop = FALSE]
rownames(data_ordered1) <- paste0("S", 1:nrow(data_ordered1))
rownames(annotation_ordered1) <- rownames(data_ordered1)

# Order rows by Structure 2 clustering
order_struct2 <- order(dataset$true_labels_2)
data_ordered2 <- data_matrix[order_struct2, ]
annotation_ordered2 <- annotation_both[order_struct2, , drop = FALSE]
rownames(data_ordered2) <- paste0("S", 1:nrow(data_ordered2))
rownames(annotation_ordered2) <- rownames(data_ordered2)

# Create heatmap ordered by Structure 1
pheatmap(data_ordered1, 
         annotation_row = annotation_ordered1,
         annotation_colors = ann_colors,
         scale = "column",
         cluster_rows = FALSE,
         cluster_cols = FALSE,
         show_rownames = FALSE,
         main = "Data Ordered by Structure 1 (Dims 1-5)",
         fontsize = 12,
         annotation_legend = TRUE,
         legend = TRUE,
         border_color = NA,
         gaps_col = c(5),  # Gap between structures
         cellwidth = 30,
         cellheight = 10)

# Create heatmap ordered by Structure 2
pheatmap(data_ordered2, 
         annotation_row = annotation_ordered2,
         annotation_colors = ann_colors,
         scale = "column",
         cluster_rows = FALSE,
         cluster_cols = FALSE,
         show_rownames = FALSE,
         main = "Data Ordered by Structure 2 (Dims 6-10)",
         fontsize = 12,
         annotation_legend = TRUE,
         legend = TRUE,
         border_color = NA,
         gaps_col = c(5),  # Gap between structures
         cellwidth = 30,
         cellheight = 10)

cat("\nKey Observation: When samples are ordered by Structure 1, the Structure 2 pattern")
cat("\nin dimensions 6-10 appears scrambled (and vice versa). This demonstrates why")
cat("\ntraditional clustering cannot capture both structures simultaneously.\n")
```

# 4. Step 1: Dimension-wise Clustering with MCLUST

We apply model-based clustering (MCLUST) to each dimension independently. This allows the Bayesian Information Criterion (BIC) to determine the optimal number of clusters for each dimension without prior knowledge.

```{r mclust_analysis}
# Initialise results storage
mclust_results <- list()
cluster_assignments <- matrix(NA, nrow = nrow(data_matrix), ncol = 10)
colnames(cluster_assignments) <- colnames(data_matrix)

# Perform MCLUST on each dimension
cat("\nApplying MCLUST to each dimension independently...\n")
cat("This allows automatic detection of cluster numbers via BIC.\n\n")

for(i in 1:10) {
  # Run Mclust on single dimension
  # G = 1:9 allows testing up to 9 clusters
  mclust_fit <- Mclust(data_matrix[, i], G = 1:9, verbose = FALSE)
  
  # Store results
  mclust_results[[paste0("Dim", i)]] <- mclust_fit
  cluster_assignments[, i] <- mclust_fit$classification
  
  # Print summary
  cat(sprintf("Dimension %d: %d clusters detected (True: %d)\n", 
              i, 
              mclust_fit$G,
              ifelse(i <= 5, 3, 4)))
}

# Summary table
optimal_summary <- data.frame(
  Dimension = 1:10,
  Detected_Clusters = sapply(mclust_results, function(x) x$G),
  True_Clusters = c(rep(3, 5), rep(4, 5)),
  Correct = sapply(mclust_results, function(x) x$G) == c(rep(3, 5), rep(4, 5))
)

cat("\nDetection accuracy:", sum(optimal_summary$Correct), "out of 10 dimensions\n")
```

```{r bic_visualisation, fig.height=6}
# Extract BIC values for visualisation
bic_data <- data.frame()
for(i in 1:10) {
  bic_values <- mclust_results[[i]]$BIC
  for(g in 1:9) {
    if(!is.na(bic_values[g])) {
      bic_data <- rbind(bic_data, 
                       data.frame(Dimension = paste0("Dim", i),
                                DimNum = i,  # Add numeric dimension for proper ordering
                                Clusters = g,
                                BIC = bic_values[g],
                                Structure = ifelse(i <= 5, "Structure 1", "Structure 2")))
    }
  }
}

# Convert Dimension to factor with correct ordering
bic_data$Dimension <- factor(bic_data$Dimension, 
                            levels = paste0("Dim", 1:10))

# Create BIC curves plot
p1 <- ggplot(bic_data, aes(x = Clusters, y = BIC, colour = Structure)) +
  geom_line(size = 1) +
  geom_point(size = 2) +
  facet_wrap(~ Dimension, scales = "free_y", ncol = 5) +
  scale_x_continuous(breaks = 1:9) +
  scale_colour_manual(values = c("Structure 1" = "#3B528B", "Structure 2" = "#86d549")) +
  labs(title = "BIC Model Selection for Dimension 7",
       subtitle = "Peak indicates optimal number of clusters",
       x = "Number of Clusters",
       y = "BIC Value") +
  theme_minimal() +
theme(
    legend.position = "bottom",
    plot.title = element_text(size = 12, hjust = 0.5, face = "bold", family = "Times"),
    plot.subtitle = element_text(size = 12, hjust = 0.5, family = "Times"),
    plot.caption = element_text(size = 12, family = "Times"),
    axis.title = element_text(size = 12, face = "bold", family = "Times"),
    axis.text = element_text(size = 10, family = "Times"),
    legend.title = element_text(size = 12, face = "bold", family = "Times"),
    legend.text = element_text(size = 10, family = "Times"),
    strip.text = element_text(size = 12, family = "Times")
  )

ggsave("1f_bicplots.pdf", plot = p1, width = 6, height = 4, units = "in")
```

# 5. Step 2: Measure Clustering Similarity with ARI

We compute the Adjusted Rand Index (ARI) between all pairs of dimension-wise clusterings. High ARI values indicate similar clustering structures.

```{r compute_ari}
# Initialise ARI matrix
n_dims <- ncol(cluster_assignments)
ari_matrix <- matrix(NA, nrow = n_dims, ncol = n_dims)
rownames(ari_matrix) <- colnames(ari_matrix) <- paste0("Dim", 1:n_dims)

# Calculate ARI for all pairs of dimensions
cat("\nComputing pairwise clustering similarity (ARI) between dimensions...\n")

for(i in 1:n_dims) {
  for(j in 1:n_dims) {
    if(i == j) {
      ari_matrix[i, j] <- 1.0
    } else {
      ari_matrix[i, j] <- adjustedRandIndex(cluster_assignments[, i], 
                                            cluster_assignments[, j])
    }
  }
}

# Calculate summary statistics
within_struct1 <- ari_matrix[1:5, 1:5][upper.tri(ari_matrix[1:5, 1:5])]
within_struct2 <- ari_matrix[6:10, 6:10][upper.tri(ari_matrix[6:10, 6:10])]
between_struct <- ari_matrix[1:5, 6:10]

cat("\nARI Summary:\n")
cat("- Within Structure 1 (dims 1-5):", round(mean(within_struct1), 3), "\n")
cat("- Within Structure 2 (dims 6-10):", round(mean(within_struct2), 3), "\n")
cat("- Between structures:", round(mean(between_struct), 3), "\n")
cat("\nThe near-zero between-structure ARI confirms the structures are independent.\n")
```

```{r visualise_ari, fig.height=8}
library(pheatmap)
library(viridis)

# Create annotation for the heatmap
annotation_ari <- data.frame(
  Structure = factor(c(rep("Structure 1", 5), rep("Structure 2", 5))),
  `Detected Clusters` = factor(as.character(optimal_summary$Detected_Clusters), levels = c("2","3","4")),
  row.names = paste0("Dim", 1:10)
)

# Define viridis-friendly annotation colors
ann_colors_ari <- list(
  Structure = c(
    "Structure 1" = "#3B528B",  # e.g., "#3B528B"
    "Structure 2" = "#86d549"          # bright greenish viridis
  ),
  `Detected Clusters` = c(
    "2" = viridis(100)[80],  # yellow-green
    "3" = viridis(100)[60],  # green
    "4" = viridis(100)[40]   # teal
  )
)

# Create the ARI heatmap
pheatmap(
  ari_matrix,
  main = "Clustering Similarity (ARI) Between Dimensions",
  annotation_col = annotation_ari,
  annotation_row = annotation_ari,
  annotation_colors = ann_colors_ari,
  color = viridis(100),
  breaks = seq(0, 1, length.out = 101),
  cluster_rows = FALSE,
  cluster_cols = FALSE,
  display_numbers = TRUE,
  number_format = "%.2f",
  number_color = "black",
  fontsize_number = 8,
  gaps_row = c(5),
  gaps_col = c(5),
  border_color = NA,
  cellwidth = 35,
  cellheight = 35,
  legend_breaks = c(0, 0.5, 1),
  fontsize = 10,             # General font size
  fontsize_row = 10,
  fontsize_col = 10,
  fontsize_main = 12,        # Title font size
  fontsize_legend = 12,      # Legend font size
  fontfamily = "Times"       # Font family
)

cat("\nThe block structure is clearly visible: high similarity within each structure,\n")
cat("near-zero similarity between structures.\n")
```

# 6. Step 3: Group Dimensions Using Hierarchical Clustering

We use Hierarchical Clustering to cluster the dimensions based on their ARI dissimilarity. The optimal number of clusters is determined by maximising the average silhouette width.

```{r MDS}
dissim_matrix <- 1 - ari_matrix
diag(dissim_matrix) <- 0  # Ensure diagonal is exactly 0

dissim_matrix2 <- dissim_matrix + abs(matrix(rnorm(ncol(data_matrix)^2, 0, 0.001), ncol(data_matrix)))

mds <- cmdscale(dissim_matrix2, k = 2)
mds <- as.data.frame(mds)
mds <- mds %>%
  mutate(group = c(rep(1, 5), rep(2, 5))) %>%
  mutate(group = as.factor(group))

# Plot MDS
p1 <- ggplot(mds, aes(x = mds[,1], y = mds[,2], color = group)) + 
  geom_point() + 
  labs(x = "MDS1",
       y = "MDS2",
       title = "2D Multi-dimensional scaling plot of dissimilarity matrix",
       color = "Structure") + 
  scale_color_manual(values = c("1" = "#3B528B", "2" = "#86d549")) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 12, face = "bold", family = "Times"),
    axis.title = element_text(size = 12, face = "bold", family = "Times"),
    axis.text = element_text(size = 10, family = "Times"),
    legend.title = element_text(size = 12, face = "bold", family = "Times"),
    legend.text = element_text(size = 10, family = "Times")
  )

print(p1)

file_path <- "1f_mdsplot_clean.pdf"
ggsave(file_path, plot = p1, width = 6, height = 4, units = "in")

mc <- mclustBootstrapLRT(mds[1:2], modelName = "EEV", nboot = 1000, level = 0.95, maxG = 1)
```

```{r pam_clustering}
# Convert ARI similarity to dissimilarity
cat("\nApplying Hierarchical clustering to group dimensions with similar clustering patterns...\n\n")

dissim_dist <- as.dist(dissim_matrix)

# Test different numbers of clusters
opt <- NbClust::NbClust(diss = dissim_dist, distance = NULL,
                        min.nc = 2, max.nc = 6, method = "complete",
                        index = "dunn")

if (opt$Best.nc[2] == Inf) {
  
  cat(sprintf("Optimal Number of Clusters:\n k = %d: Dunn Index = %.3f\n", 
                opt$Best.nc[1], opt$Best.nc[2]))
  cat("\nInfinite Dunn Index -> Switching to Silhouette Index\n")
  
  k_values <- 2:6
  silhouette_widths <- numeric(length(k_values))
  hclust_results <- list()
  
  for(i in seq_along(k_values)) {
    k <- k_values[i]
    hclust_fit <- hclust(dissim_dist, method = "complete")
    clusters <- cutree(hclust_fit, k = k)
    hclust_results[[as.character(k)]] <- clusters
    silhouette_widths[i] <- mean(silhouette(clusters, dist = dissim_dist)[, "sil_width"])
    
    cat(sprintf("\n k = %d: Average silhouette width = %.3f", 
                k, silhouette_widths[i]))
  }
  
  # Find optimal number of clusters
  optimal_k <- k_values[which.max(silhouette_widths)]
  optimal_hclust <- hclust_results[[as.character(optimal_k)]]
  
  cat("\nOptimal number of dimension groups:", optimal_k, "\n")
  
} else {
  
  for (i in 1:length(opt$All.Index)) {
    
    k <- names(opt$All.index)[i]
    val <- opt$All.index[[i]]
    cat(sprintf("k = %s: Dunn Index = %.3f\n", 
                k, val))
    
  }
  
  # Find optimal number of clusters
  optimal_k <- opt$Best.nc[1]
  optimal_hclust <- opt$Best.partition
  
}


```

```{r visualise_pam, fig.height=5}
# Create silhouette width plot
silhouette_df <- data.frame(
  k = k_values,
  avg_silhouette = silhouette_widths
)

p1 <- ggplot(silhouette_df, aes(x = k_values, y = avg_silhouette)) +
  geom_line(size = 1.2, colour = "#B8DE29") +
  geom_point(size = 3, colour = "#B8DE29") +
  geom_vline(xintercept = optimal_k, linetype = "dashed", colour = "#2A788E", size = 1) +
  geom_text(aes(x = optimal_k, y = max(avg_silhouette) * 0.95, 
                label = paste("Optimal k =", optimal_k)),
            hjust = -0.1, colour = "#2A788E", size = 4) +
  scale_x_continuous(breaks = k_values) +
  labs(title = "Selecting Optimal Number of Dimension Groups",
       subtitle = "Higher silhouette width indicates better clustering quality",
       x = "Number of Clusters (k)",
       y = "Average Silhouette Width") +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 12, hjust = 0.5, face = "bold", family = "Times"),
    plot.subtitle = element_text(size = 12, hjust = 0.5, family = "Times"),
    plot.caption = element_text(size = 12, family = "Times"),
    axis.title = element_text(size = 12, face = "bold", family = "Times"),
    axis.text = element_text(size = 10, family = "Times"),
    legend.title = element_text(size = 12, face = "bold", family = "Times"),
    legend.text = element_text(size = 10, family = "Times"),
    strip.text = element_text(size = 12, family = "Times")
  )

print(p1)

ggsave("1f_sil.pdf", plot = p1, width = 6, height = 4, units = "in")

# Print cluster membership
cat("\nHierarchical Clustering Assignments:\n")
cluster_assignments_hclust <- optimal_hclust
names(cluster_assignments_hclust) <- paste0("Dim", 1:10)

for(i in 1:optimal_k) {
  members <- names(cluster_assignments_hclust)[cluster_assignments_hclust == i]
  cat(sprintf("Group %d: %s\n", i, paste(members, collapse = ", ")))
}
```

# 7. Step 4: Validate Structure Recovery

We assess whether Hierarchical Clustering has successfully separated the two known structures.

```{r validate_recovery}
# Check separation of structures
struct1_clusters <- unique(cluster_assignments_hclust[1:5])
struct2_clusters <- unique(cluster_assignments_hclust[6:10])

cat("\n\nValidation Results:\n")
cat("Structure 1 dimensions assigned to HClust cluster(s):", 
    paste(sort(struct1_clusters), collapse = ", "), "\n")
cat("Structure 2 dimensions assigned to HClust cluster(s):", 
    paste(sort(struct2_clusters), collapse = ", "), "\n")

# Calculate recovery ARI
true_structure_labels <- c(rep(1, 5), rep(2, 5))
recovery_ari <- adjustedRandIndex(cluster_assignments_hclust, true_structure_labels)

cat("\nRecovery Performance (ARI):", round(recovery_ari, 3), "\n")

if(recovery_ari == 1) {
  cat("\n✓ Perfect recovery! The two structures have been completely separated.\n")
} else if(recovery_ari > 0.8) {
  cat("\n✓ Excellent recovery! The structures are well separated.\n")
} else {
  cat("\n✗ Partial recovery. Some dimensions may be misclassified.\n")
}
```

```{r final_visualisation, fig.height=8}
# Reorder ARI matrix by Hiearchical clusters for final visualisation
cluster_order <- order(cluster_assignments_hclust)
ari_reordered <- ari_matrix[cluster_order, cluster_order]

# Create comprehensive annotation
annotation_final <- data.frame(
  `HClust Group` = factor(cluster_assignments_hclust[cluster_order]),
  `True Structure` = factor(c(rep("Structure 1", 5), rep("Structure 2", 5))[cluster_order]),
  `Detected k` = optimal_summary$Detected_Clusters[cluster_order],
  row.names = names(cluster_assignments_hclust)[cluster_order]
)

# Define colours
ann_colors_final <- list(
  `HClust Group` = c("1" = "#FF6B6B", "2" = "#4ECDC4"),
  `True Structure` = c("Structure 1" = "#377EB8", "Structure 2" = "#E41A1C"),
  `Detected k` = c("2" = "#FEE08B", "3" = "#FDAE61", "4" = "#F46D43")
)

# Create final heatmap
pheatmap(ari_reordered,
         main = "Final Result: Dimensions Grouped by Clustering Similarity",
         annotation_col = annotation_final,
         annotation_row = annotation_final,
         annotation_colors = ann_colors_final,
         color = colorRampPalette(c("white", "lightyellow", "orange", "red", "darkred"))(100),
         breaks = seq(0, 1, length.out = 101),
         cluster_rows = FALSE,
         cluster_cols = FALSE,
         display_numbers = TRUE,
         number_format = "%.2f",
         number_color = "black",
         fontsize_number = 8,
         border_color = NA,
         cellwidth = 35,
         cellheight = 35,
         gaps_row = cumsum(table(cluster_assignments_hclust[cluster_order]))[1:(optimal_k-1)],
         gaps_col = cumsum(table(cluster_assignments_hclust[cluster_order]))[1:(optimal_k-1)],
         fontsize = 10)
```

# 8. Step 5: Consensus Clustering with COCA

Now that we've identified which dimensions belong to each structure using Hiearchical clustering, we can use COCA (Cluster-Of-Clusters Analysis) to create a single consensus clustering for each discovered structure. This combines the information from multiple dimension-wise clusterings into a robust summary.

```{r coca_setup}
# Load COCA library
library(coca)

# Set seed for reproducibility
set.seed(42)

# Define fixed version of consensusCluster
consensusClusterPDWK <- function (data = NULL, K = 2, B = 100, pItem = 0.8, clMethod = "hclust", 
          dist = "euclidean", hclustMethod = "complete", sparseKmeansPenalty = NULL, 
          maxIterKM = 1000) 
{
  # Set seed for reproducibility within this function
  set.seed(42)
  
  containsFactors <- 0
  if (!is.null(data)) {
    N <- dim(data)[1]
    P <- dim(data)[2]
    for (i in seq_len(P)) {
      containsFactors <- as.numeric(is.factor(data[, i])) + 
        containsFactors
    }
  }
  else if (is.double(dist)) {
    N <- dim(dist)[1]
  }
  else {
    stop("If the data matrix is not provided, `dist` must be a symmetric\n        matrix of type double providing the distances between each pair of\n        observations.")
  }
  dataIndices <- seq_len(N)
  coClusteringMatrix <- indicatorMatrix <- matrix(0, N, N)
  for (b in seq_len(B)) {
    # Set seed for each iteration to ensure reproducibility
    set.seed(42 + b)
    
    items <- sample(N, ceiling(N * pItem), replace = FALSE)
    nUniqueDataPoints <- 0
    if (!is.null(data)) {
      uniqueData <- unique(data[items, ])
      nUniqueDataPoints <- nrow(uniqueData)
    }
    if (nUniqueDataPoints >= K | is.null(data)) {
      if (clMethod == "pam" | containsFactors) {
        if (is.double(dist) & isSymmetric(dist)) {
          distances <- stats::as.dist(dist[items, items])
        }
        else if (dist == "cor") {
          distances <- stats::as.dist(1 - stats::cor(t(data[items, 
          ])))
        }
        else if (dist == "binary") {
          distances <- stats::dist(data[items, ], method = dist)
        }
        else if (dist == "gower") {
          distances <- cluster::daisy(data[items, ], 
                                      metric = "gower")
        }
        else {
          stop("Distance not recognized. If method is `pam`, distance\n                must be one of `cor`, `binary`, `gower` or the symmetric\n                matrix of distances.")
        }
        cl <- cluster::pam(distances, K)$clustering
      }
      else if (clMethod == "kmeans" & !is.null(data)) {
        cl <- stats::kmeans(data[items, ], K, iter.max = maxIterKM, 
                            nstart = 20)$cluster
      }
      else if (clMethod == "sparse-kmeans" & !is.null(data)) {
        if (is.null(sparseKmeansPenalty)) 
          sparseKmeansPenalty = sqrt(P)
        cat("sparseKmeansPenalty", sparseKmeansPenalty, 
            "\n")
        cl <- sparcl::KMeansSparseCluster(data[items, 
        ], K, wbounds = sparseKmeansPenalty)[[1]]$Cs
      }
      else if (clMethod == "hclust" | clMethod == "sparse-hclust") {
        if (dist == "pearson" | dist == "spearman") {
          pearsonCor <- stats::cor(t(data[items, ]), 
                                   method = dist)
          distances <- stats::as.dist(1 - pearsonCor)
        }
        else if (is.double(dist)) {
          distances <- stats::as.dist(dist[items, items])
        }
        else {
          distances <- stats::dist(data[items, ], method = dist)
        }
        if (clMethod == "hclust") {
          hClustering <- stats::hclust(distances, method = hclustMethod)
        }
        else {
          hClustering <- sparcl::HierarchicalSparseCluster(dists = as.matrix(distances), 
                                                           method = "average", wbound = 10)$hc
        }
        cl <- stats::cutree(hClustering, K)
      }
      else {
        stop("Clustering algorithm name not recognised. Please choose\n                     one of `kmeans`, `hclust`, `pam`, `sparse-kmeans`,\n                     `sparse-hclust`.")
      }
      indicatorMatrix <- indicatorMatrix + crossprod(t(as.numeric(dataIndices %in% 
                                                                    items)))
      for (k in seq_len(K)) {
        coClusteringMatrix[items, items] <- coClusteringMatrix[items, 
                                                               items] + crossprod(t(as.numeric(cl == k)))
      }
    }
  }
  if (!sum(indicatorMatrix) == 0) {
    consensusMatrix <- coClusteringMatrix/indicatorMatrix
  }
  else {
    consensusMatrix <- indicatorMatrix
    warning(paste("Consensus matrix is empty for K =", K, 
                  "because there are\n                      less than", 
                  K, "distinct data points", sep = ""))
  }
  return(consensusMatrix)
}

# Save original function and replace with fixed version
consensusClusterOriginal <- coca::consensusCluster
assignInNamespace("consensusCluster", consensusClusterPDWK, ns="coca")

# Extract dimension groups as identified by HClust clustering
# We found that HClust identified optimal_k groups
n_groups <- optimal_k
dimension_groups <- list()

for(g in 1:n_groups) {
  dimension_groups[[g]] <- which(cluster_assignments_hclust == g)
}

cat("HClust identified", n_groups, "dimension groups:\n")
for(g in 1:n_groups) {
  cat("Group", g, ":", paste0("Dim", dimension_groups[[g]]), "\n")
}
```

## 8.1 Build MOC and Apply COCA for Each Discovered Group

```{r coca_analysis}
# Store results for each group
coca_results <- list()
consensus_clusterings <- list()
moc_list <- list()

for(g in 1:n_groups) {
  cat("\n\n=== Analyzing Group", g, "===\n")
  group_dims <- dimension_groups[[g]]
  cat("Dimensions in this group:", paste0("Dim", group_dims), "\n")
  
  # Set seed for reproducibility in buildMOC
  set.seed(42 + g)
  
  # Create list of data matrices for this group (one per dimension)
  group_data_list <- list()
  for(i in 1:length(group_dims)) {
    # Each "dataset" is just one column (dimension) of our data
    group_data_list[[i]] <- data_matrix[, group_dims[i], drop = FALSE]
  }
  
  # Get the cluster numbers for each dimension
  K_values <- sapply(group_dims, function(d) mclust_results[[paste0("Dim", d)]]$G)
  cat("Number of clusters in each dimension:", paste(K_values, collapse = ", "), "\n")
  
  # Build MOC using the coca package function
  # Note: buildMOC expects each dataset to cluster independently
  moc_output <- buildMOC(group_data_list, 
                        M = length(group_dims),
                        K = K_values,
                        methods = rep("kmeans", length(group_dims)),
                        distances = rep("euclidean", length(group_dims)))
  
  # Extract the MOC
  MOC <- moc_output$moc
  moc_list[[g]] <- MOC
  
  cat("MOC dimensions:", nrow(MOC), "clusters ×", ncol(MOC), "observations\n")
  
  # Apply COCA and let it determine the optimal K
  # Use the mode of detected clusters to set maxK
  mode_k <- as.numeric(names(sort(table(K_values), decreasing = TRUE)[1]))
  maxK_value <- max(6, mode_k + 2)  # Allow searching up to mode + 2 or 6, whichever is larger
  
  cat("Applying COCA with automatic K selection (maxK =", maxK_value, ")\n")
  
  # Apply COCA with automatic K selection
  coca_result <- coca(MOC, 
                     K = NULL,  # Let COCA determine optimal K
                     maxK = maxK_value,
                     B = 100,  # Use same B as in consensusCluster
                     pItem = 0.8,
                     choiceKmethod = "silhouette",
                     ccClMethod = "kmeans",
                     returnAllMatrices = TRUE,  # Get all consensus matrices
                     verbose = TRUE)
  
  coca_results[[g]] <- coca_result
  consensus_clusterings[[g]] <- coca_result$clusterLabels
  
  cat("\nCOCA selected K =", coca_result$K, "for Group", g, "\n")
  cat("Final cluster sizes:", table(consensus_clusterings[[g]]), "\n")
  
  # Show the consensus matrix quality
  consensus_mat <- coca_result$consensusMatrix
  mean_consensus <- mean(consensus_mat[upper.tri(consensus_mat)])
  cat("Mean consensus value:", round(mean_consensus, 3), "\n")
  
  # Validation if we know the true structure
  if(all(group_dims <= 5)) {
    true_labels_for_group <- dataset$true_labels_1
    true_structure <- "Structure 1"
    true_k <- 3
  } else if(all(group_dims > 5)) {
    true_labels_for_group <- dataset$true_labels_2  
    true_structure <- "Structure 2"
    true_k <- 4
  } else {
    true_labels_for_group <- NULL
    true_structure <- "Mixed"
    true_k <- NA
  }
  
  if(!is.null(true_labels_for_group)) {
    ari_with_truth <- adjustedRandIndex(consensus_clusterings[[g]], true_labels_for_group)
    cat("ARI with true", true_structure, "labels:", round(ari_with_truth, 3), "\n")
    cat("True K =", true_k, ", COCA selected K =", coca_result$K, "\n")
  }
}

# Restore original consensusCluster function
assignInNamespace("consensusCluster", consensusClusterOriginal, ns="coca")
cat("\nRestored original consensusCluster function\n")
```

## 8.2 Visualize Matrix Of Clusters (MOC)

```{r moc_visualization, fig.height=10}
# Visualize the MOC for each group
for(g in 1:n_groups) {
  MOC <- moc_list[[g]]
  
  # Check dimensions
  cat("\nGroup", g, "MOC dimensions:", nrow(MOC), "rows ×", ncol(MOC), "columns\n")

  MOC_display <- MOC
  
  # Order by consensus clustering
  order_idx <- order(consensus_clusterings[[g]])
  MOC_ordered <- MOC_display[order_idx, ]
    
  
  # Create row labels
  if(nrow(MOC) == n_samples) {
    # If we transposed, create cluster labels
    K_values <- sapply(dimension_groups[[g]], function(d) mclust_results[[paste0("Dim", d)]]$G)
    col_labels <- paste0("Dim", rep(dimension_groups[[g]], times = K_values), 
                        "_Cluster", unlist(lapply(K_values, function(k) 1:k)))
  } else {
    col_labels <- colnames(MOC_ordered)
  }
  
  file = paste0("leap_MOC_Group",g,".png")
  
  # Plot MOC
  pheatmap(MOC_ordered,
           main = paste0("Group ", g, " - Matrix Of Clusters (MOC)\n",
                        "Binary matrix: rows = observations, columns = clusters"),
           color = c("white", "black"),
           breaks = c(-0.5, 0.5, 1.5),
           cluster_rows = FALSE,  # Allow row clustering
           cluster_cols = TRUE,
           show_rownames = FALSE,
           show_colnames = TRUE,
           border_color = "grey80",
           legend = FALSE,
         cellwidth = 1.5,
           cellheight = 0.5,
        fontsize_row = 10,
        fontsize_col = 10,
        fontsize_main = 12,        
        fontsize_legend = 10,      
        fontfamily = "Times",
        filename = file,
        width = 6,
        height = 8)
  
  # Summary based on actual orientation
  cat("\nGroup", g, "MOC summary:\n")
  cat("- Dimensions included:", paste0("Dim", dimension_groups[[g]]), "\n")
  if(nrow(MOC) == n_samples) {
    cat("- Observations (rows):", nrow(MOC), "\n")
    cat("- Total clusters (columns):", ncol(MOC), "\n")
  } else {
    cat("- Total clusters (rows):", nrow(MOC), "\n")
    cat("- Observations (columns):", ncol(MOC), "\n")
  }
  cat("- Sparsity:", round(sum(MOC == 0) / length(MOC) * 100, 1), "% zeros\n")
}

cat("\n\nNote: buildMOC appears to return the MOC with observations in rows and")
cat("\nclusters in columns, which is the transpose of the notation in the COCA paper.\n")
```

## 8.3 Visualize COCA Consensus Matrices

```{r coca_visualization, fig.height=8}
# Function to plot consensus matrix with discovered clusters
plot_coca_consensus <- function(coca_result, group_num, group_dims) {
  # Extract consensus matrix and cluster labels from coca result
  consensus_mat <- coca_result$consensusMatrix
  clustering <- coca_result$clusterLabels
  
  # Order by consensus clustering
  order_idx <- order(clustering)
  consensus_ordered <- consensus_mat[order_idx, order_idx]
  
  # Create title
  title <- paste0("Group ", group_num, " (", paste0("Dim", group_dims, collapse=", "), 
                  "): COCA Consensus Matrix (K=", coca_result$K, ")")
  
  # Generate colors based on number of clusters
  n_clusters <- length(unique(clustering))
  if(n_clusters <= 3) {
    my_colors <- c("#E41A1C", "#377EB8", "#4DAF4A")[1:n_clusters]
  } else if(n_clusters <= 8) {
    my_colors <- RColorBrewer::brewer.pal(n_clusters, "Set2")
  } else {
    my_colors <- colorRampPalette(RColorBrewer::brewer.pal(8, "Set2"))(n_clusters)
  }
  
  file = paste0("1f_COCAheatmap_Group",group_num,".png")
  
  # Simple heatmap without row/column annotations to avoid the error
  pheatmap(consensus_ordered,
           main = title,
           color = viridis(100),
           breaks = seq(0, 1, length.out = 101),
           cluster_rows = FALSE,
           cluster_cols = FALSE,
           show_rownames = FALSE,
           show_colnames = FALSE,
         annotation_legend = TRUE,
         legend = TRUE,
         border_color = NA,
         cellwidth = 10,
           cellheight = 8,
        fontsize_row = 10,
        fontsize_col = 10,
        fontsize_main = 12,        
        fontsize_legend = 10,      
        fontfamily = "Times",
        filename = file,
        width = 6,
        height = 8)
  
  # Add text to show the clustering
  cat("\nGroup", group_num, "consensus clustering:", 
      paste(clustering[order_idx], collapse=" "), "\n")
  cat("Number of clusters:", n_clusters, "\n")
  cat("Cluster sizes:", paste(table(clustering), collapse=", "), "\n")
}

# Plot consensus matrix for each group
for(g in 1:n_groups) {
  plot_coca_consensus(coca_results[[g]], g, dimension_groups[[g]])
}

cat("\nThe COCA consensus matrices show how often pairs of observations were")
cat("\nclustered together across the consensus clustering iterations.\n")
cat("Dark red indicates high consensus (often clustered together),")
cat("\nwhite indicates low consensus (rarely clustered together).\n")
```

## 8.4 Compare Dimension-wise vs COCA Consensus

```{r compare_clusterings}
# Create comparison table
comparison_df <- data.frame()

for(g in 1:n_groups) {
  group_dims <- dimension_groups[[g]]
  
  # Compare each dimension's clustering with the COCA consensus
  for(d in group_dims) {
    ari_with_consensus <- adjustedRandIndex(cluster_assignments[, d], 
                                           consensus_clusterings[[g]])
    
    comparison_df <- rbind(comparison_df, data.frame(
      Group = g,
      Dimension = paste0("Dim", d),
      ARI_with_Consensus = ari_with_consensus,
      Detected_k = mclust_results[[paste0("Dim", d)]]$G
    ))
  }
  
  # Add consensus result
  comparison_df <- rbind(comparison_df, data.frame(
    Group = g,
    Dimension = "COCA Consensus",
    ARI_with_Consensus = 1.0,
    Detected_k = length(unique(consensus_clusterings[[g]]))
  ))
}

# Display comparison
knitr::kable(comparison_df, digits = 3,
             caption = "Agreement between Individual Dimensions and COCA Consensus",
             col.names = c("Group", "Method", "ARI with Consensus", "k"))

cat("\nThe table shows how well individual dimension clusterings agree with")
cat("\nthe COCA consensus within each discovered group.\n")
```

## 8.5 Final Visualization: All Consensus Clusterings

```{r final_visualization, fig.height=8, fig.width = 12}
library(viridisLite)

# Create combined annotation showing all consensus clusterings
combined_annotation <- data.frame(row.names = paste0("S", 1:n_samples))

# Add consensus clustering for each group
for (g in 1:n_groups) {
  col_name <- paste0("Group", g, "_Consensus")
  combined_annotation[[col_name]] <- factor(consensus_clusterings[[g]])
}

# Add true structure labels (if available)
if (exists("dataset") && !is.null(dataset$true_labels_1)) {
  combined_annotation$True_Struct1 <- factor(dataset$true_labels_1)
  combined_annotation$True_Struct2 <- factor(dataset$true_labels_2)
}

# Define annotation colors
ann_colors <- list()

# Assign viridis palettes for each GroupX_Consensus
for (g in 1:n_groups) {
  col_name <- paste0("Group", g, "_Consensus")
  levels_vec <- sort(unique(combined_annotation[[col_name]]))
  ann_colors[[col_name]] <- setNames(viridis(length(levels_vec), option = "D"), as.character(levels_vec))
}

# Assign similar but distinguishable palettes for True_Struct1/2
if ("True_Struct1" %in% colnames(combined_annotation)) {
  levels_1 <- sort(unique(combined_annotation$True_Struct1))
  levels_2 <- sort(unique(combined_annotation$True_Struct2))
  
  ann_colors$True_Struct1 <- setNames(viridis(length(levels_1), option = "C"), as.character(levels_1))
  ann_colors$True_Struct2 <- setNames(viridis(length(levels_2), option = "B"), as.character(levels_2))
}

# Rename annotation columns for display in legend
colnames(combined_annotation)[colnames(combined_annotation) == "Group1_Consensus"] <- "Structure 1\nConsensus Cluster\nLabels"
colnames(combined_annotation)[colnames(combined_annotation) == "Group2_Consensus"] <- "Structure 2\nConsensus Cluster\nLabels"
colnames(combined_annotation)[colnames(combined_annotation) == "True_Struct1"] <- "True Cluster\nLabels for\nStructure 1"
colnames(combined_annotation)[colnames(combined_annotation) == "True_Struct2"] <- "True Cluster\nLabels for\nStructure 2"

names(ann_colors)[names(ann_colors) == "Group1_Consensus"] <- "Structure 1\nConsensus Cluster\nLabels"
names(ann_colors)[names(ann_colors) == "Group2_Consensus"] <- "Structure 2\nConsensus Cluster Labels"
names(ann_colors)[names(ann_colors) == "True_Struct1"] <- "True Cluster\nLabels for\nStructure 1"
names(ann_colors)[names(ann_colors) == "True_Struct2"] <- "True Cluster\nLabels for\nStructure 2"


# Order by consensus clustering
order_idx <- order(consensus_clusterings[[1]])
if (n_groups > 1) {
  order_idx <- order(consensus_clusterings[[1]], consensus_clusterings[[2]])
}

data_ordered_final <- data_matrix[order_idx, ]
annotation_ordered_final <- combined_annotation[order_idx, , drop = FALSE]
rownames(data_ordered_final) <- rownames(annotation_ordered_final)

# Plot heatmap
pheatmap(data_ordered_final,
         annotation_row = annotation_ordered_final,
         annotation_colors = ann_colors,
         scale = "column",
         cluster_rows = FALSE,
         cluster_cols = FALSE,
         show_rownames = FALSE,
         main = "Final Result: Data with COCA Consensus Clusterings for Each Discovered Group",
```
         
```{r final_visualization, fig.height=8, fig.width = 6}
library(viridisLite)

# Create combined annotation showing all consensus clusterings
combined_annotation <- data.frame(row.names = paste0("S", 1:n_samples))

# Add consensus clustering for each group
for (g in 1:n_groups) {
  col_name <- paste0("Group", g, "_Consensus")
  combined_annotation[[col_name]] <- factor(consensus_clusterings[[g]])
}

# Add true structure labels (if available)
if (exists("dataset") && !is.null(dataset$true_labels_1)) {
  combined_annotation$True_Struct1 <- factor(dataset$true_labels_1)
  combined_annotation$True_Struct2 <- factor(dataset$true_labels_2)
}

# Define annotation colors
ann_colors <- list()

# Assign viridis palettes for each GroupX_Consensus
for (g in 1:n_groups) {
  col_name <- paste0("Group", g, "_Consensus")
  levels_vec <- sort(unique(combined_annotation[[col_name]]))
  ann_colors[[col_name]] <- setNames(viridis(length(levels_vec), option = "D"), as.character(levels_vec))
}

# Assign similar but distinguishable palettes for True_Struct1/2
if ("True_Struct1" %in% colnames(combined_annotation)) {
  levels_1 <- sort(unique(combined_annotation$True_Struct1))
  levels_2 <- sort(unique(combined_annotation$True_Struct2))
  
  ann_colors$True_Struct1 <- setNames(viridis(length(levels_1), option = "C"), as.character(levels_1))
  ann_colors$True_Struct2 <- setNames(viridis(length(levels_2), option = "B"), as.character(levels_2))
}

# Rename annotation columns for display in legend
colnames(combined_annotation)[colnames(combined_annotation) == "Group1_Consensus"] <- "S1 Consensus"
colnames(combined_annotation)[colnames(combined_annotation) == "Group2_Consensus"] <- "S2 Consensus"
colnames(combined_annotation)[colnames(combined_annotation) == "True_Struct1"] <- "S1 True"
colnames(combined_annotation)[colnames(combined_annotation) == "True_Struct2"] <- "S2 True"

names(ann_colors)[names(ann_colors) == "Group1_Consensus"] <- "S1 Consensus"
names(ann_colors)[names(ann_colors) == "Group2_Consensus"] <- "S2 Consensus"
names(ann_colors)[names(ann_colors) == "True_Struct1"] <- "S1 True"
names(ann_colors)[names(ann_colors) == "True_Struct2"] <- "S2 True"


# Order by consensus clustering
order_idx <- order(consensus_clusterings[[1]])
if (n_groups > 1) {
  order_idx <- order(consensus_clusterings[[1]], consensus_clusterings[[2]])
}

data_ordered_final <- data_matrix[order_idx, ]
annotation_ordered_final <- combined_annotation[order_idx, , drop = FALSE]
rownames(data_ordered_final) <- rownames(annotation_ordered_final)

# Plot heatmap
pheatmap(data_ordered_final,
         annotation_row = annotation_ordered_final,
         annotation_colors = ann_colors,
         scale = "column",
         cluster_rows = FALSE,
         cluster_cols = FALSE,
         show_rownames = FALSE,
         main = "Final Result: Data with COCA Consensus\nClusterings for Each Discovered Group",
         annotation_legend = TRUE,
         legend = TRUE,
         border_color = NA,
         gaps_col = cumsum(sapply(dimension_groups, length))[1:(n_groups - 1)],
         cellwidth = 24,
           cellheight = 8,
        fontsize_row = 10,
        fontsize_col = 10,
        fontsize_main = 12,        
        fontsize_legend = 10,      
        fontfamily = "Times",
        filename = "1f_finalheatmap.png",
        width = 6,
        height = 8)
```
# 9. Summary: Complete Analysis Pipeline

Our complete methodology now includes:

1. **Dimension-wise Clustering**: Applied MCLUST to each dimension independently
2. **Similarity Assessment**: Computed pairwise ARI between dimension clusterings  
3. **Dimension Grouping**: Used HClust to identify dimensions with similar patterns
4. **Structure Validation**: Confirmed successful separation of clustering structures
5. **Consensus Clustering**: Applied COCA to get a single robust clustering for each discovered group

## Final Results

```{r summary_results}
cat("Summary of Results:\n")
cat("==================\n\n")

cat("1. HClust discovered", n_groups, "dimension groups:\n")
for(g in 1:n_groups) {
  cat("   Group", g, ":", paste0("Dim", dimension_groups[[g]]), "\n")
}

cat("\n2. COCA consensus clustering for each group:\n")
for(g in 1:n_groups) {
  n_clusters <- length(unique(consensus_clusterings[[g]]))
  cat("   Group", g, ": Found", n_clusters, "clusters\n")
}

# If we know the true structure (for validation)
if(n_groups == 2 && all(dimension_groups[[1]] <= 5) && all(dimension_groups[[2]] > 5)) {
  cat("\n3. Validation against known structure:\n")
  ari_1 <- adjustedRandIndex(consensus_clusterings[[1]], dataset$true_labels_1)
  ari_2 <- adjustedRandIndex(consensus_clusterings[[2]], dataset$true_labels_2)
  cat("   Group 1 (Structure 1): ARI =", round(ari_1, 3), "\n")
  cat("   Group 2 (Structure 2): ARI =", round(ari_2, 3), "\n")
}
```

## Benefits of the Complete Pipeline

- **Discovery**: Automatically identifies multiple clustering structures without prior knowledge
- **Interpretability**: Clear identification of which dimensions contribute to each structure
- **Robustness**: COCA consensus leverages agreement across multiple dimension-wise views
- **Actionability**: Provides a single clustering solution for each discovered structure

## Next Steps

To understand how this approach performs under more challenging conditions:
- See `Example1g.Rmd` for a comprehensive simulation study
- The simulation varies cluster separation from 6σ down to 0.5σ
- Results show at what point cluster overlap prevents structure recovery

## Conclusion

By combining dimension grouping with COCA consensus clustering, we achieve both structure discovery and robust cluster assignment. This provides a complete solution for analyzing data with multiple independent clustering structures.
