---
title: 'Simulation Study: MDS Thresholding for the Presence of Views'
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = TRUE, message = TRUE, warning = TRUE)
```

# Executive Summary

This simulation study extends our analysis to investigate how bootstrapped model-based clustering of the low-dimensional representation of the dissimilarity matrix, from Multi-dimensional Scaling, performs in determining the presence of multiple views. We test whether the approach is accurate for 1 structure (our baseline). 2, 3, 4, and 5 independent structures, all embedded in 100-dimensional data.

**Key Questions:** 1. How does the thresholding method work with differing numbers of independent structures? 2. Is there a minimum number of structures that can reliably assess the presence of views? 

# Implementation 

## 1. Methods

### 1.1 Experimental Design

We use data divided equally among k structures:
**p=10**
- **k=2**: 5 dimensions per structure (baseline)
- **k=3**: 3-4 dimensions per structure
- **k=4**: 2-3 dimensions per structure  
- **k=5**: 2 dimensions per structure
**p=100**
- **k=2**: 50 dimensions per structure (baseline)
- **k=3**: 33-34 dimensions per structure
- **k=4**: 25 dimensions per structure  
- **k=5**: 20 dimensions per structure
**p=500**
- **k=2**: 250 dimensions per structure (baseline)
- **k=3**: 166-167 dimensions per structure
- **k=4**: 125 dimensions per structure  
- **k=5**: 100 dimensions per structure

Each structure has a different number of clusters (3, 4, 5, 6, or 7) to ensure distinguishability.

### Experimental Grid:
- **Number of structures**: 1, 2, 3, 4, 5
- **Dimensions**: 10, 100, 500
- **Sample sizes**: 250, 500, 1000
- **Separation factors**: 2.5σ
- **Replications**: 25
- **Parallel cores**: 8

Total: 45 conditions × 25 replications = 1124 runs

## 2. Implementation

```{r libraries}
# Load required libraries
library(mvtnorm)      # For multivariate normal data generation
library(cluster)      # For PAM clustering
library(mclust)       # For model-based clustering
library(otrimle)      # For otrimle
library(ggplot2)      # For visualisation
library(dplyr)        # For data manipulation
library(tidyr)        # For data reshaping
library(viridis)      # For colour scales
library(gridExtra)    # For arranging plots
library(foreach)      # For parallel loops
library(doParallel)   # For parallel backend
library(parallel)     # For makeCluster and stopCluster
library(R.utils)      # For mclust timeout

# Set seed for reproducibility
set.seed(42)
```

```{r parallel_setup}
# Note: Parallel cluster will be created within the simulation chunk
# to ensure proper scoping in R Markdown environment

cat("Parallel processing configuration:\n")
cat("- Total cores available:", detectCores(), "\n")
cat("- Cores to be used:", detectCores() - 2, "\n")
cat("- Cores kept free for system:", 2, "\n\n")
```

```{r cache_setup}
# Set up caching directory
cache_dir <- "~/Documents/MPhil Population Health Sciences/Dissertation/GMM Simulation Study/mds_threshold_cache"
if (!dir.exists(cache_dir)) {
  dir.create(cache_dir)
  cat("Created cache directory:", cache_dir, "\n")
} else {
  cat("Using existing cache directory:", cache_dir, "\n")
}

# Cache functions
get_cache_filename <- function(n_structures, p_dims, n_samples, separation_factor) {
  filename <- sprintf("results_k%d_d%d_n%d_sep%.1f.rds", n_structures, p_dims, n_samples, separation_factor)
  file.path(cache_dir, filename)
}

check_condition_cache <- function(n_structures, p_dims, n_samples, separation_factor) {
  cache_file <- get_cache_filename(n_structures, p_dims, n_samples, separation_factor)
  file.exists(cache_file)
}

save_condition_results <- function(results, n_structures, p_dims, n_samples, separation_factor) {
  cache_file <- get_cache_filename(n_structures, p_dims, n_samples, separation_factor)
  saveRDS(results, cache_file)
}

load_condition_results <- function(n_structures, p_dims, n_samples, separation_factor) {
  cache_file <- get_cache_filename(n_structures, p_dims, n_samples, separation_factor)
  if (file.exists(cache_file)) {
    return(readRDS(cache_file))
  } else {
    return(NULL)
  }
}

get_all_results_cache_filename <- function() {
  file.path(cache_dir, "all_results_full.rds")
}
```

```{r parameters}
# Experimental parameters
n_structures_values <- c(1, 2, 3, 4, 5)
p_dims <- c(10, 100, 500)  
sample_sizes <- c(250, 500, 1000)
separation_factors <- c(2.5)
n_replications <- 25

# Fixed parameters
cluster_variance <- 0.3

# Clusters per structure (different for each to ensure distinguishability)
clusters_per_structure <- c(3, 4, 5, 6, 7)  # Will use first k of these

# Total conditions
n_conditions <- length(p_dims) * length(n_structures_values) * length(sample_sizes) * length(separation_factors)

cat("\nExperimental design:\n")
cat("- Number of structures (k):", paste(n_structures_values, collapse=", "), "\n")
cat("- Total dimensions:", paste(p_dims, collapse=", "), "\n")
cat("- Sample sizes:", paste(sample_sizes, collapse=", "), "\n")
cat("- Separation factors: ", paste(separation_factors, "σ", sep="", collapse=", "), "\n")
cat("- Total conditions:", n_conditions, "\n")
cat("- Replications per condition:", n_replications, "\n")
cat("- Total simulation runs:", n_conditions * n_replications, "\n")
```

```{r data_generation_functions}
# Generate cluster means for k structures
generate_cluster_means_k <- function(separation_factor, cluster_variance, n_structures, dims_list) {
  base_separation <- separation_factor * sqrt(cluster_variance)
  
  all_means <- list()
  
  for (k in 1:n_structures) {
    n_clusters <- clusters_per_structure[k]
    n_dims <- dims_list[k]
    
    # Generate means for this structure
    structure_means <- list()
    for (c in 1:n_clusters) {
      # Create distinct patterns for each cluster
      if (c <= 2) {
        # First two clusters: opposite sides
        multiplier <- ifelse(c == 1, -1, 1)
        means <- runif(n_dims, 0.8, 1.5) * multiplier * base_separation
      } else {
        # Additional clusters: random positions
        means <- runif(n_dims, -0.7, 0.7) * base_separation
        # Add some structure to avoid complete overlap
        means <- means + (c - 2.5) * 0.3 * base_separation
      }
      structure_means[[paste0("cluster_", c)]] <- means
    }
    
    all_means[[paste0("structure_", k)]] <- structure_means
  }
  
  return(all_means)
}

# Generate dataset with k structures
generate_dataset_k <- function(p_dim, separation_factor, cluster_variance, n_samples, n_structures, dims_list) {
  
  # Generate cluster assignments for each structure
  cluster_assignments <- list()
  for (k in 1:n_structures) {
    n_clusters <- clusters_per_structure[k]
    # Use different proportions for each structure
    props <- diff(c(0, sort(runif(n_clusters-1)), 1))
    cluster_assignments[[k]] <- sample(1:n_clusters, n_samples, replace = TRUE, prob = props)
  }
  
  # Get cluster means
  all_means <- generate_cluster_means_k(separation_factor, cluster_variance, n_structures, dims_list)
  
  # Initialize data matrix
  data_matrix <- matrix(NA, nrow = n_samples, ncol = p_dim)
  
  # Generate data for each structure
  col_start <- 1
  for (k in 1:n_structures) {
    n_dims <- dims_list[k]
    col_end <- col_start + n_dims - 1
    
    structure_means <- all_means[[paste0("structure_", k)]]
    
    # Generate data for this structure
    for (i in 1:n_samples) {
      cluster_id <- cluster_assignments[[k]][i]
      cluster_mean <- structure_means[[paste0("cluster_", cluster_id)]]
      cluster_cov <- diag(rep(cluster_variance, n_dims))
      
      data_matrix[i, col_start:col_end] <- mvtnorm::rmvnorm(1, mean = cluster_mean, sigma = cluster_cov)
    }
    
    col_start <- col_end + 1
  }
  
  return(list(
    data = data_matrix,
    true_assignments = cluster_assignments,
    dims_list = dims_list
  ))
}
```
```{r analysis_functions}
# Analyze dataset with k structures
analyze_dataset_k <- function(data_matrix, n_structures, separation_factor, n_samples, dims_list) {
  n_dims <- ncol(data_matrix)

  # Step 1: MCLUST on each dimension
  cluster_assignments <- matrix(NA, nrow = nrow(data_matrix), ncol = n_dims)
  detected_clusters <- numeric(n_dims)

  for (i in 1:n_dims) {
    mclust_fit <- mclust::Mclust(data_matrix[, i], G = 1:9, verbose = FALSE)
    cluster_assignments[, i] <- mclust_fit$classification
    detected_clusters[i] <- mclust_fit$G
  }

  # Step 2: Compute ARI matrix
  ari_matrix <- matrix(1, nrow = n_dims, ncol = n_dims)
  for (i in 1:(n_dims - 1)) {
    for (j in (i + 1):n_dims) {
      ari_val <- mclust::adjustedRandIndex(cluster_assignments[, i], cluster_assignments[, j])
      ari_matrix[i, j] <- ari_val
      ari_matrix[j, i] <- ari_val
    }
  }

  # Step 3: Compute dissimilarity matrix with small noise
  dis_mat <- (1 - ari_matrix) + abs(matrix(rnorm(n_dims * n_dims, 0, 0.001), n_dims, n_dims))
  dist_mat <- as.dist(dis_mat)

  # Step 4: MDS
  mds_result <- cmdscale(dist_mat, k = 2, eig = TRUE)
  mds <- as.data.frame(mds_result$points)

  # Step 5: Model loop
  models <- c("EEE", "EEI", "EEV", "EII", "EVE", "EVI", "EVV", "OTRIMLE")
  results <- data.frame()

  for (model in models) {
    # Default values for all cases
    start_time <- Sys.time()
    rt <- NA
    stat <- "ERROR"
    dns <- "NA"
    method_label <- if (model == "OTRIMLE") "OTRIMLE" else paste("mclustBootstrapLRT:", model)

    if (model != "OTRIMLE") {
      # Try MCLUST model
      test <- tryCatch({
        R.utils::withTimeout({
          mclust::mclustBootstrapLRT(mds, modelName = model, nboot = 1000, level = 0.95, maxG = 1)
        }, timeout = 10800, onTimeout = "error")
      }, error = function(e) {
        message(sprintf("❌ MCLUST model %s failed: %s", model, e$message))
        NULL
      })

      end_time <- Sys.time()

      if (!is.null(test) && is.list(test) && "p.value" %in% names(test)) {
        p <- round(test$p.value, 4)
        dns <- ifelse(!is.na(p) && p < 0.05, ">1", "1")
        stat <- paste("p-value:", p)
        rt <- as.numeric(difftime(end_time, start_time, units = "secs"))
      } else {
        stat <- "p-value: NA"
      }

    } else {
      # Try OTRIMLE
      model_otrimle <- tryCatch({
        R.utils::withTimeout({
          otrimle::otrimleg(mds, G = 1:2)
        }, timeout = 10800, onTimeout = "error")
      }, error = function(e) {
        message(sprintf("❌ OTRIMLE failed: %s", e$message))
        NULL
      })

      end_time <- Sys.time()

      if (!is.null(model_otrimle) && is.list(model_otrimle) && "desncrit" %in% names(model_otrimle)) {
        min_denscrit <- min(model_otrimle$denscrit, na.rm = TRUE)
        best_G <- which.min(model_otrimle$denscript)
        dns <- ifelse(best_G == 2, ">1", "1")
        stat <- paste("min(Q):", round(min_denscrit, 4))
        rt <- as.numeric(difftime(end_time, start_time, units = "secs"))
      } else {
        stat <- "min(BIC): NA"
      }
    }

    # Add result row
    cat_val <- ifelse(n_structures == 1, "1", ">1")
    results <- rbind(results, data.frame(
      tns = n_structures,
      cat = cat_val,
      p_dim = n_dims,
      sep_fact = separation_factor,
      n_samples = n_samples,
      method = method_label,
      stat = stat,
      dns = dns,
      correct = (dns == cat_val),
      rt = rt,
      stringsAsFactors = FALSE
    ))
  }

  cat("✅ Number of rows in results:", nrow(results), "\n")
  return(results)
}


# Wrapper function for parallel execution
run_single_replication <- function(p_dim, rep, n_structures, separation_factor, n_samples, dims_list) {
  # Generate dataset
  dataset <- generate_dataset_k(p_dim, separation_factor, cluster_variance, n_samples, n_structures, dims_list)
  
  # Analyze dataset
  analysis_result <- analyze_dataset_k(dataset$data, n_structures, separation_factor, n_samples, dims_list)
  
  # Add metadata to result
  analysis_result$replication <- rep
  
  print(paste("Replication", rep, "- nrow(analysis_result):", nrow(analysis_result)))
  cat("In run_single_replication, rows in analysis_result:", nrow(analysis_result), "\n")
  
  return(analysis_result)
}

# Safe wrapper function
safe_run_single_replication <- function(p_dim, rep, n_structures, separation_factor, n_samples, dims_list, max_time_sec = 10800) {
    result <- NULL
    tryCatch({
        result <- withTimeout({
            run_single_replication(
                p_dim = p_dim,
                rep = rep,
                n_structures = n_structures,
                separation_factor = separation_factor,
                n_samples = n_samples,
                dims_list = dims_list
            )
        }, timeout = max_time_sec, onTimeout = "error")
        
        # DEBUG: Print class and structure of the result immediately after running
        cat(sprintf("Replication %d returned an object of class: %s\n", rep, class(result)))
        if (!is.list(result)) {
            cat("Warning: Result is not a list! Here's a snapshot:\n")
            print(str(result))
        }
        
    }, error = function(e) {
        warning(sprintf("Replication %d failed or timed out: %s", rep, e$message))
        result <<- NULL
    })
    return(result)
}

safe_run_single_model <- function(p_dim, rep, n_structures, separation_factor, n_samples, dims_list, model) {
  result <- NULL
  tryCatch({
    # Generate data (same across models for this rep)
    dataset <- generate_dataset_k(p_dim, separation_factor, cluster_variance, n_samples, n_structures, dims_list)
    data_matrix <- dataset$data

    # Run adjusted Rand Index clustering steps (used for dissimilarity matrix)
    cluster_assignments <- matrix(NA, nrow = nrow(data_matrix), ncol = ncol(data_matrix))
    detected_clusters <- numeric(ncol(data_matrix))
    
    for (i in 1:ncol(data_matrix)) {
      mclust_fit <- mclust::Mclust(data_matrix[, i], G = 1:9, verbose = FALSE)
      cluster_assignments[, i] <- mclust_fit$classification
      detected_clusters[i] <- mclust_fit$G
    }

    ari_matrix <- matrix(1, ncol(data_matrix), ncol(data_matrix))
    for (i in 1:(ncol(data_matrix)-1)) {
      for (j in (i+1):ncol(data_matrix)) {
        ari_val <- mclust::adjustedRandIndex(cluster_assignments[, i], cluster_assignments[, j])
        ari_matrix[i, j] <- ari_val
        ari_matrix[j, i] <- ari_val
      }
    }

    dis_mat <- (1 - ari_matrix) + abs(matrix(rnorm(ncol(data_matrix)^2, 0, 0.001), ncol(data_matrix)))
    dist_mat <- as.dist(dis_mat)
    mds_result <- cmdscale(dist_mat, k = 2, eig = TRUE)
    mds <- as.data.frame(mds_result$points)

    # Run the selected model on the MDS-reduced data
    result <- analyze_single_model(model, mds, n_structures, separation_factor, p_dim, n_samples)

  }, error = function(e) {
    warning(sprintf("Model %s failed for replication %d: %s", model, rep, e$message))
  })

  # Add metadata
  result$p_dims <- p_dim
  result$n_structures <- n_structures
  result$n_samples <- n_samples
  result$replication <- rep
  
  return(result)
}

analyze_single_model <- function(model, mds, n_structures, separation_factor, p_dim, n_samples) {
  start_time <- Sys.time()
  end_time <- NA
  rt <- NA
  stat <- "ERROR"
  dns <- "NA"
  method_label <- if (toupper(model) == "OTRIMLE") "OTRIMLE" else paste("mclustBootstrapLRT:", model)

  if (toupper(model) != "OTRIMLE") {
    tryCatch({
      test <- R.utils::withTimeout({
        mclust::mclustBootstrapLRT(
          mds,
          modelName = model,
          nboot = 1000,
          level = 0.95,
          maxG = 1
        )
      }, timeout = 10800, onTimeout = "error")

      end_time <- Sys.time()

      pval <- if (!is.null(test$p.value)) round(test$p.value, 4) else NA
      dns <- ifelse(!is.na(pval) && pval < 0.05, ">1", "1")
      stat <- paste("p-value:", pval)
    }, error = function(e) {
      end_time <<- Sys.time()
      stat <<- paste("ERROR:", e$message)
      dns <<- "NA"
    })
  } else {
    tryCatch({
      model_otrimle <- R.utils::withTimeout({
        otrimle::otrimleg(mds, G = 1:2)
      }, timeout = 10800, onTimeout = "error")

      end_time <- Sys.time()

      if (!is.list(model_otrimle) || !"denscrit" %in% names(model_otrimle)) {
        stop("OTRIMLE result is malformed")
      }

      best_G <- which.min(model_otrimle$denscrit)
      min_denscrit <- min(model_otrimle$denscrit, na.rm = TRUE)

      dns <- ifelse(best_G == 2, ">1", "1")
      stat <- paste("min(Q):", round(min_denscrit, 4))
    }, error = function(e) {
      end_time <<- Sys.time()
      stat <<- paste("ERROR:", e$message)
      dns <<- "NA"
    })
  }

  # Finalize runtime computation
  if (!is.na(end_time)) {
    rt <- as.numeric(difftime(end_time, start_time, units = "secs"))
  }

  # Correct label
  cat_val <- ifelse(n_structures == 1, "1", ">1")

  # Return as a consistent data frame
  return(data.frame(
    tns = n_structures,
    cat = cat_val,
    method = method_label,
    stat = stat,
    dns = dns,
    correct = (dns == cat_val),
    rt = rt,
    sep_fact = separation_factor,
    p_dims = p_dim,
    n_samples = n_samples,
    stringsAsFactors = FALSE
  ))
}
```

# 3. Simulation

```{r run_simulation, cache=TRUE}
library(foreach)
library(doParallel)
library(dplyr)

cat("Starting simulation chunk\n")

# Number of cores for parallel (leave 2 free)
n_cores <- max(1, parallel::detectCores() - 2)

# Initialize cluster
cl <- makeCluster(n_cores)
registerDoParallel(cl)

models <- c(
      "mclustBootstrapLRT: EEE", "mclustBootstrapLRT: EEI", "mclustBootstrapLRT: EEV",
      "mclustBootstrapLRT: EII", "mclustBootstrapLRT: EVE", "mclustBootstrapLRT: EVI",
      "mclustBootstrapLRT: EVV", "OTRIMLE"
    )

# Build full grid of conditions
conditions_grid <- expand.grid(
  n_structures = n_structures_values,
  p_dim = p_dims,
  n_samples = sample_sizes,
  separation_factor = separation_factors
)

# Check which conditions already have cached results
conditions_grid$cached <- mapply(
  check_condition_cache,
  conditions_grid$n_structures,
  conditions_grid$p_dim,
  conditions_grid$n_samples,
  conditions_grid$separation_factor
)

# Select only uncached conditions to run
to_run <- subset(conditions_grid, cached == FALSE)

if (nrow(to_run) == 0) {
  message("✅ All simulation results are already cached.")
} else {
  message("🚀 Running uncached simulation conditions.")

  foreach(i = 1:nrow(to_run),
          .packages = c("mvtnorm", "cluster", "mclust", "NbClust", "otrimle", "dplyr", "R.utils","foreach"),
          .errorhandling = "pass",
          .verbose = TRUE) %dopar% {

    cond <- to_run[i, ]

    # Dynamically compute dims_list correctly
    dims_list <- rep(floor(cond$p_dim / cond$n_structures), cond$n_structures)
    remainder <- cond$p_dim %% cond$n_structures
    if (remainder > 0) {
      dims_list[1:remainder] <- dims_list[1:remainder] + 1
    }

    # Models to run per your original list
    models <- c(
      "mclustBootstrapLRT: EEE", "mclustBootstrapLRT: EEI", "mclustBootstrapLRT: EEV",
      "mclustBootstrapLRT: EII", "mclustBootstrapLRT: EVE", "mclustBootstrapLRT: EVI",
      "mclustBootstrapLRT: EVV", "OTRIMLE"
    )

    # Load cached results if any (in case partial)
    existing_results <- NULL
    if (check_condition_cache(cond$n_structures, cond$p_dim, cond$n_samples, cond$separation_factor)) {
      existing_results <- load_condition_results(cond$n_structures, cond$p_dim, cond$n_samples, cond$separation_factor)
    }

    all_results <- if (!is.null(existing_results)) existing_results else data.frame()

    rep_results <- lapply(1:n_replications, function(rep) {
      existing_methods <- if (nrow(all_results) > 0) {
        subset(all_results, replication == rep & !is.na(dns))$method
      } else {
        character(0)
      }
    
      missing_models <- setdiff(models, existing_methods)
      print(paste("Replication", rep, "- missing models:", paste(missing_models, collapse = ", ")))
    
      if (length(missing_models) > 0) {
        partial_results <- foreach(
          model = missing_models,
          .combine = rbind,
          .packages = c("mvtnorm", "mclust", "cluster", "otrimle", "R.utils")
        ) %do% {
          safe_run_single_model(
            p_dim = cond$p_dim,
            rep = rep,
            n_structures = cond$n_structures,
            separation_factor = cond$separation_factor,
            n_samples = cond$n_samples,
            dims_list = dims_list,
            model = model
          )
        }
        partial_results
      } else {
        NULL
      }
    })


    rep_results <- Filter(Negate(is.null), rep_results)
    combined_results <- rbind(all_results, do.call(rbind, rep_results))
    
    save_condition_results(combined_results,
                           cond$n_structures,
                           cond$p_dim,
                           cond$n_samples,
                           cond$separation_factor)

  }
}

# Stop cluster
stopCluster(cl)

# Optionally: load and flatten all cached results as in first script
all_results_files <- list.files(cache_dir, pattern = "results_k.*\\.rds$", full.names = TRUE)

all_conditions_results <- lapply(all_results_files, function(file) {
  result_df <- tryCatch(readRDS(file), error = function(e) NULL)
  if (!is.null(result_df) && is.data.frame(result_df)) {
    result_df
  } else if (is.list(result_df)) {
    do.call(rbind, result_df)
  } else {
    NULL
  }
})

all_conditions_results <- Filter(Negate(is.null), all_conditions_results)
all_results_df <- do.call(rbind, all_conditions_results)

message("✅ Simulation complete. Total rows: ", nrow(all_results_df))
```

# 4. Results

```{r load-all-results, echo=FALSE, message=FALSE, warning=FALSE}
all_results <- readRDS("~/Documents/MPhil Population Health Sciences/Dissertation/GMM Simulation Study/mds_threshold_cache/all_results.rds")
```

```{r full-results, cache=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(knitr)
library(ggplot2)
library(stringr)

results_clean <- all_results %>% 
  mutate(dns = ifelse(dns == "NA", NA, dns),
         correct = ifelse(is.na(dns), NA, correct), 
                                        runtime = rt) %>%
  select(-rt)

results_clean$method <- stringr::str_remove(results_clean$method, "mclustBootstrapLRT: ")

results_clean <- results_clean %>%
  mutate(
    truth_binary = case_when(
      is.na(cat) ~ NA_integer_,
      cat == "1"  ~ 0L,
      cat == ">1" ~ 1L,
      TRUE ~ NA_integer_  
    ),
    detected_binary = case_when(
      is.na(dns) ~ NA_integer_,
      dns == "1"  ~ 0L,
      dns == ">1" ~ 1L,
      TRUE ~ NA_integer_
    ),
    
    TP = if_else(is.na(truth_binary) | is.na(detected_binary), NA,
                 truth_binary == 1 & detected_binary == 1),
    TN = if_else(is.na(truth_binary) | is.na(detected_binary), NA,
                 truth_binary == 0 & detected_binary == 0),
    FP = if_else(is.na(truth_binary) | is.na(detected_binary), NA,
                 truth_binary == 0 & detected_binary == 1),
    FN = if_else(is.na(truth_binary) | is.na(detected_binary), NA,
                 truth_binary == 1 & detected_binary == 0)
  )


# === Overall summary per method (Exact Clustering Performance) ===
overall_summary <- results_clean %>%
  group_by(method) %>%
  summarise(
    Exact_Correct_Rate = mean(correct, na.rm = TRUE),
    Mean_Runtime_sec = mean(runtime, na.rm = TRUE),
    NA_Count = sum(is.na(runtime)),
    Total_Runs = n()
  ) %>%
  arrange(desc(Exact_Correct_Rate))

# === Confusion matrix metrics per method (Binary Classification) ===
confusion_summary <- results_clean %>%
  group_by(method) %>%
  summarise(
    TP_Total = sum(TP, na.rm = TRUE),
    TN_Total = sum(TN, na.rm = TRUE),
    FP_Total = sum(FP, na.rm = TRUE),
    FN_Total = sum(FN, na.rm = TRUE),
    NA_count = sum(is.na(TP) | is.na(TN) | is.na(FP) | is.na(FN)),
    Accuracy = (TP_Total + TN_Total) / (TP_Total + TN_Total + FP_Total + FN_Total),
    Sensitivity_TPR = TP_Total / (TP_Total + FN_Total),
    Specificity_TNR = TN_Total / (TN_Total + FP_Total),
    FPR = FP_Total / (FP_Total + TN_Total),
    FNR = FN_Total / (FN_Total + TP_Total)
  ) %>%
  arrange(desc(Accuracy))

# === Confusion matrix metrics per method and sample size (Binary Classification) ===
confusion_summary_s <- results_clean %>%
  group_by(method, n_samples) %>%
  summarise(
    TP_Total = sum(TP, na.rm = TRUE),
    TN_Total = sum(TN, na.rm = TRUE),
    FP_Total = sum(FP, na.rm = TRUE),
    FN_Total = sum(FN, na.rm = TRUE),
    NA_count = sum(is.na(TP) | is.na(TN) | is.na(FP) | is.na(FN)),
    Accuracy = (TP_Total + TN_Total) / (TP_Total + TN_Total + FP_Total + FN_Total),
    Sensitivity_TPR = TP_Total / (TP_Total + FN_Total),
    Specificity_TNR = TN_Total / (TN_Total + FP_Total),
    FPR = FP_Total / (FP_Total + TN_Total),
    FNR = FN_Total / (FN_Total + TP_Total)
  ) %>%
  arrange(desc(Accuracy))

# === Confusion matrix metrics per method and structures (Binary Classification) ===
confusion_summary_st <- results_clean %>%
  group_by(method, n_structures) %>%
  summarise(
    TP_Total = sum(TP, na.rm = TRUE),
    TN_Total = sum(TN, na.rm = TRUE),
    FP_Total = sum(FP, na.rm = TRUE),
    FN_Total = sum(FN, na.rm = TRUE),
    NA_count = sum(is.na(TP) | is.na(TN) | is.na(FP) | is.na(FN)),
    Accuracy = (TP_Total + TN_Total) / (TP_Total + TN_Total + FP_Total + FN_Total),
    Sensitivity_TPR = TP_Total / (TP_Total + FN_Total),
    Specificity_TNR = TN_Total / (TN_Total + FP_Total),
    FPR = FP_Total / (FP_Total + TN_Total),
    FNR = FN_Total / (FN_Total + TP_Total)
  ) %>%
  arrange(desc(Accuracy))

# === Breakdown by true number of structures (Exact Clustering Performance) ===
breakdown_structures <- results_clean %>%
  group_by(method, n_structures) %>%
  summarise(
    Exact_Correct_Rate = mean(correct, na.rm = TRUE),
    Mean_Runtime_sec = mean(runtime, na.rm = TRUE),
        NA_Count = sum(is.na(runtime)),
    Total_Runs = n()
  ) %>%
  ungroup() %>% 
  arrange(n_structures, desc(Exact_Correct_Rate))

# === Breakdown by sample size (Exact Clustering Performance) ===
breakdown_samples <- results_clean %>%
  group_by(method, n_samples) %>%
  summarise(
    Exact_Correct_Rate = mean(correct, na.rm = TRUE),
    Mean_Runtime_sec = mean(runtime, na.rm = TRUE),
    NA_Count = sum(is.na(runtime)),
    Total_Runs = n()
  ) %>%
  ungroup()

# === Missing data analysis 
missing_by_method <- results_clean %>%
  group_by(method) %>%
  summarise(
    Missing_Count = sum(is.na(runtime)),
    Total = n(),
    Missing_Rate = Missing_Count / Total
  ) %>%
  arrange(desc(Missing_Rate))

missing_by_structures <- results_clean %>%
  group_by(method, n_structures) %>%
  summarise(
    Missing_Count = sum(is.na(runtime)),
    Total = n(),
    Missing_Rate = Missing_Count / Total
  ) %>%
  ungroup()

missing_by_samples <- results_clean %>%
  group_by(method, n_samples) %>%
  summarise(
    Missing_Count = sum(is.na(runtime)),
    Total = n(),
    Missing_Rate = Missing_Count / Total
  ) %>%
  ungroup()



# === Print overall summaries ===
cat("## Overall Exact Clustering Performance Summary\n")
kable(overall_summary, digits = 3, caption = "Exact Correctness, Runtime (sec), and NA Counts by Method")

cat("\n\n## Binary Classification Metrics by Method\n")
kable(confusion_summary, digits = 3, caption = "Binary Classification Metrics by Method")

cat("\n\n## Binary Classification Metrics by Method and Number of Structures \n")
kable(confusion_summary_st, digits = 3, caption = "Binary Classification Metrics by Method and Number of Structures")

cat("\n\n## Binary Classification Metrics by Method and Sample Size\n")
kable(confusion_summary_s, digits = 3, caption = "Binary Classification Metrics by Method and Sample Size")

cat("\n\n## Exact Clustering Correctness Rate by True Number of Structures\n")
kable(breakdown_structures, digits = 3, caption = "Exact Correctness and Runtime by True Number of Structures")

cat("\n\n## Exact Clustering Correctness Rate by Sample Size\n")
kable(breakdown_samples, digits = 3, caption = "Exact Correctness and Runtime by Sample Size")

cat("\n\n## Missing Data Summary by Method\n")
kable(missing_by_method, digits = 3, caption = "Missing Data by Method")

cat("\n\n## Missing Data Summary by Number of Structures\n")
kable(missing_by_structures, digits = 3, caption = "Missing Data by Number of Structures")

cat("\n\n## Missing Data Summary by Sample Size\n")
kable(missing_by_samples, digits = 3, caption = "Missing Data by Sample Size")



# === Plots ===

p1 <- ggplot(breakdown_structures, 
             aes(x = factor(n_structures), 
                 y = Exact_Correct_Rate, 
                 color = method,
                 shape = NA_Count > 0)) +  # Map shape based on condition
  geom_point(size = 3, alpha = 1) + 
  geom_line(aes(group = method)) +
  labs(
    title = "Accuracy of Detecting Multiple Views by True Number of Structures, \nAccounting for Missing Data",
    x = "True Number of Structures",
    y = "Accuracy"
  ) +
  theme_minimal() + 
  scale_color_viridis_d(name = "Covariance Structure", option = "viridis") +
  scale_shape_manual(
    name = "Missing Data",  # Optional legend title
    values = c(`FALSE` = 16, `TRUE` = 17)  # 16 = solid circle, 17 = triangle
  ) +
  theme(
    plot.title = element_text(size = 12, face = "bold", family = "Times"),
    plot.subtitle = element_text(size = 12, family = "Times"),
    plot.caption = element_text(size = 12, family = "Times"),
    axis.title = element_text(size = 12, face = "bold", family = "Times"),
    axis.text = element_text(size = 10, family = "Times"),
    legend.title = element_text(size = 12, face = "bold", family = "Times"),
    legend.text = element_text(size = 10, family = "Times"),
    strip.text = element_text(size = 12, family = "Times")
  )

print(p1)
ggsave("1l_correctness.pdf", plot = p1, width = 6, height = 4, units = "in")

p4 <- ggplot(breakdown_samples,
             aes(x = factor(n_samples),
                 y = Exact_Correct_Rate,
                 color = method,
                 group = method)) +  # connect points by method
  geom_point(size = 3, alpha = 1) +
  geom_line() +
  labs(
    title = "Accuracy of Detecting Multiple Views by Sample Size",
    x = "Sample Size",
    y = "Accuracy"
  ) +
  theme_minimal() +
  scale_color_viridis_d(name = "Covariance Structure", option = "viridis") +
  theme(
    plot.title = element_text(size = 12, face = "bold", family = "Times"),
    axis.title = element_text(size = 12, face = "bold", family = "Times"),
    axis.text = element_text(size = 10, family = "Times"),
    legend.title = element_text(size = 12, face = "bold", family = "Times"),
    legend.text = element_text(size = 10, family = "Times")
  )

print(p4)
ggsave("1l_accuracy_by_sample_size.pdf", plot = p4, width = 6, height = 4, units = "in")



confusion_long <- confusion_summary %>%
  select(method, Sensitivity_TPR, Specificity_TNR) %>%
  pivot_longer(cols = c(Sensitivity_TPR, Specificity_TNR), names_to = "Metric", values_to = "Value")

p2 <- ggplot(confusion_long, aes(x = method, y = Value, fill = Metric)) +
  geom_col(position = "dodge") +
  labs(title = "Sensitivity (TPR) and Specificity (TNR) by Method", y = "Value", x = "Method") +
  theme_minimal() +
  coord_flip()

print(p2)

fn_fp_long <- confusion_summary %>%
  select(method, FPR, FNR) %>%
  pivot_longer(cols = c(FPR, FNR), names_to = "Metric", values_to = "Value")

p3 <- ggplot(fn_fp_long, aes(x = method, y = Value, fill = Metric)) +
  geom_col(position = "dodge") +
  labs(title = "False Positive Rate and False Negative Rate by Method", y = "Value", x = "Method") +
  theme_minimal() +
  coord_flip()

print(p3)
```